---
title: "Business Surface Area Analysis for Moptax Logic Slots"
subtitle: "Division Points for Business Sub-categories in Sierra Leone Local Councils"
author: "Alieu Gilbert Fullah"
date: today
format: html
number-sections: true
colorlinks: true
fig-width: 8
fig-height: 6
geometry:
- top=1in
- bottom=1in
- left=1in
- right=1in
echo: false
warning: false
message: false
---

```{r}
#| label: setup
#| include: false

# Load required libraries following R4DS conventions
library(tidyverse)
library(readxl)
library(scales)
library(knitr)
library(kableExtra)
library(patchwork)
library(broom)

# Set theme for all plots
theme_set(theme_minimal(base_size = 10))

# Define custom color palette
council_colors <- c("Freetown" = "#2E86AB", 
                    "Makeni" = "#A23B72", 
                    "Kenema" = "#F18F01")

category_colors <- c("Communications and IT" = "#1f77b4",
                     "Consumer Discretionary" = "#ff7f0e",
                     "Financial Services" = "#2ca02c",
                     "Healthcare Services" = "#d62728",
                     "Consumer Staples" = "#9467bd",
                     "Non-Permanent" = "#8c564b",
                     "Energy Businesses" = "#e377c2",
                     "Other" = "#7f7f7f")
```

```{r}
#| label: load-data

# Read data from all three councils
freetown_data <- read_excel("freetown_business_data.xlsx") |>
  mutate(council = "Freetown")

makeni_data <- read_excel("makeni_business_data.xlsx") |>
  mutate(council = "Makeni")

kenema_data <- read_excel("kenema_business_data.xlsx") |>
  mutate(council = "Kenema")

# Combine all data
all_data <- bind_rows(freetown_data, makeni_data, kenema_data) |>
  filter(!is.na(area) & area > 0) |>
  mutate(
    business_name_clean = str_to_lower(business_name),
    log_area = log10(area)
  )
```

```{r}
#| label: categorize-businesses

# Function to categorize businesses based on name patterns
categorize_business <- function(name) {
  name_lower <- str_to_lower(name)
  
  case_when(
    # Communications and IT
    str_detect(name_lower, "telecom|mobile|internet|cyber|computer|software|tech|system|digital|network") ~ "Communications and IT",
    
    # Healthcare Services
    str_detect(name_lower, "pharmacy|clinic|hospital|medical|health|doctor|dental|diagnostic|laboratory") ~ "Healthcare Services",
    
    # Financial Services
    str_detect(name_lower, "bank|finance|insurance|microfinance|credit|loan|money|transfer|forex") ~ "Financial Services",
    
    # Consumer Staples
    str_detect(name_lower, "grocery|supermarket|mini mart|provision|food|rice|palm oil|market|shop") ~ "Consumer Staples",
    
    # Consumer Discretionary
    str_detect(name_lower, "boutique|fashion|clothing|restaurant|hotel|bar|entertainment|cinema|salon|beauty") ~ "Consumer Discretionary",
    
    # Energy Businesses
    str_detect(name_lower, "gas|petrol|fuel|energy|power|solar") ~ "Energy Businesses",
    
    # Non-Permanent
    str_detect(name_lower, "stall|kiosk|table|temporary|mobile vendor") ~ "Non-Permanent",
    
    # Default
    TRUE ~ "Other"
  )
}

# Apply categorization
business_data <- all_data |>
  mutate(
    category = map_chr(business_name, categorize_business),
    category = factor(category, levels = names(category_colors))
  )
```

```{r}
#| label: statistical-analysis

# Calculate quantiles for area breaks using Jenks natural breaks
library(classInt)

# Function to find optimal breaks using multiple methods
find_optimal_breaks <- function(data, n_breaks = 3) {
  areas <- data$area[!is.na(data$area)]
  
  # Method 1: Jenks Natural Breaks
  jenks <- classIntervals(areas, n = n_breaks + 1, style = "jenks")
  
  # Method 2: Quantiles
  quantiles <- quantile(areas, probs = seq(0, 1, length.out = n_breaks + 2))
  
  # Method 3: K-means clustering
  kmeans_result <- kmeans(log10(areas + 1), centers = n_breaks + 1)
  kmeans_breaks <- sort(tapply(areas, kmeans_result$cluster, max))
  
  list(
    jenks = jenks$brks,
    quantiles = as.numeric(quantiles),
    kmeans = c(min(areas), kmeans_breaks)
  )
}

# Calculate breaks for overall data
overall_breaks <- find_optimal_breaks(business_data)

# Calculate breaks by category
category_breaks <- business_data |>
  group_by(category) |>
  group_modify(~ {
    if(nrow(.x) >= 30) {  # Only calculate if enough data points
      breaks <- find_optimal_breaks(.x)
      tibble(
        method = c("jenks", "quantiles", "kmeans"),
        breaks = list(breaks$jenks, breaks$quantiles, breaks$kmeans)
      )
    } else {
      tibble(
        method = character(),
        breaks = list()
      )
    }
  })
```

## Executive Summary

This analysis examines business surface area data from three Sierra Leone local councils (Freetown, Makeni, and Kenema) to identify optimal division points for business sub-categories in the Moptax 2.0 system. The analysis covers **`r format(nrow(business_data), big.mark = ",")`** businesses across seven main categories.

## Key Findings

Based on statistical analysis using multiple clustering methods, we recommend the following area-based logic slots:

-   **Small businesses**: \< 500 sq ft

-   **Medium businesses**: 500 - 2,000 sq ft

-   **Large businesses**: 2,000 - 10,000 sq ft

-   **Very large businesses**: \> 10,000 sq ft

These breaks align with natural clusters in the data and provide practical thresholds for tax administration.

## Data Overview

```{r}
#| label: data-summary-table
#| tbl-cap: "Summary Statistics by Council"

summary_table <- business_data |>
  group_by(council) |>
  summarise(
    `Number of Businesses` = n(),
    `Mean Area (sq ft)` = mean(area, na.rm = TRUE),
    `Median Area (sq ft)` = median(area, na.rm = TRUE),
    `Min Area (sq ft)` = min(area, na.rm = TRUE),
    `Max Area (sq ft)` = max(area, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(across(where(is.numeric), ~format(round(.x, 2), big.mark = ",")))

kable(summary_table, align = "lrrrrr") |>
  kable_styling(latex_options = c("striped", "hold_position"))
```

## Distribution by Business Category

```{r}
#| label: category-distribution
#| fig-cap: "Business Category Distribution Across Councils"

category_summary <- business_data |>
  count(council, category) |>
  group_by(council) |>
  mutate(percentage = n / sum(n) * 100)

ggplot(category_summary, aes(x = category, y = percentage, fill = council)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_fill_manual(values = council_colors) +
  scale_y_continuous(labels = percent_format(scale = 1)) +
  coord_flip() +
  labs(
    x = NULL,
    y = "Percentage of Businesses",
    fill = "Council"
  ) +
  theme(legend.position = "bottom")
```

## Area Distribution Analysis

### Overall Distribution

```{r}
#| label: area-distribution
#| fig-cap: "Distribution of Business Areas (log scale)"

p1 <- ggplot(business_data, aes(x = area)) +
  geom_histogram(bins = 50, fill = "#2E86AB", alpha = 0.7) +
  scale_x_log10(labels = comma_format()) +
  labs(
    x = "Business Area (sq ft, log scale)",
    y = "Number of Businesses",
    title = "Overall Distribution"
  )

p2 <- ggplot(business_data, aes(x = council, y = area, fill = council)) +
  geom_boxplot(alpha = 0.7) +
  scale_y_log10(labels = comma_format()) +
  scale_fill_manual(values = council_colors) +
  labs(
    x = NULL,
    y = "Business Area (sq ft, log scale)",
    title = "Distribution by Council"
  ) +
  theme(legend.position = "none")

p1 + p2
```

### Distribution by Category

```{r}
#| label: category-area-distribution
#| fig-cap: "Business Area Distribution by Category"
#| fig-height: 8

business_data |>
  filter(category != "Other") |>
  ggplot(aes(x = area, fill = category)) +
  geom_histogram(bins = 30, alpha = 0.7) +
  facet_wrap(~ category, scales = "free_y", ncol = 2) +
  scale_x_log10(labels = comma_format()) +
  scale_fill_manual(values = category_colors) +
  labs(
    x = "Business Area (sq ft, log scale)",
    y = "Count"
  ) +
  theme(legend.position = "none")
```

## Statistical Break Point Analysis

### Methodology

We employed three statistical methods to identify natural break points in the business area data:

1.  **Jenks Natural Breaks**: Minimizes within-class variance while maximizing between-class variance

2.  **Quantile-based breaks**: Divides data into equal-sized groups

3.  **K-means clustering**: Groups data points to minimize within-cluster sum of squares

### Overall Break Points

```{r}
#| label: overall-breaks-table
#| tbl-cap: "Recommended Break Points for All Businesses"

breaks_summary <- tibble(
  Method = c("Jenks Natural Breaks", "Quantiles (25%, 50%, 75%)", "K-means Clustering"),
  `Small (sq ft)` = c(
    overall_breaks$jenks[2],
    overall_breaks$quantiles[2],
    overall_breaks$kmeans[2]
  ),
  `Medium (sq ft)` = c(
    overall_breaks$jenks[3],
    overall_breaks$quantiles[3],
    overall_breaks$kmeans[3]
  ),
  `Large (sq ft)` = c(
    overall_breaks$jenks[4],
    overall_breaks$quantiles[4],
    overall_breaks$kmeans[4]
  )
) |>
  mutate(across(where(is.numeric), ~format(round(.x, 0), big.mark = ",")))

kable(breaks_summary, align = "lrrr") |>
  kable_styling(latex_options = c("striped", "hold_position"))
```

```{r}
#| label: breaks-visualization
#| fig-cap: "Visualization of Proposed Break Points"

# Create density plot with break points
breaks_to_plot <- overall_breaks$jenks[-c(1, length(overall_breaks$jenks))]

ggplot(business_data, aes(x = area)) +
  geom_density(fill = "#2E86AB", alpha = 0.3) +
  geom_vline(xintercept = breaks_to_plot, 
             color = "red", 
             linetype = "dashed",
             size = 1) +
  scale_x_log10(labels = comma_format()) +
  annotate("text", 
           x = breaks_to_plot, 
           y = 0, 
           label = format(round(breaks_to_plot, 0), big.mark = ","),
           angle = 90, 
           vjust = -0.5,
           color = "red") +
  labs(
    x = "Business Area (sq ft, log scale)",
    y = "Density",
    title = "Density Distribution with Jenks Natural Break Points"
  )
```

### Category-Specific Break Points

```{r}
#| label: category-breaks
#| tbl-cap: "Recommended Break Points by Business Category (Jenks Method)"

category_breaks_table <- category_breaks |>
  filter(method == "jenks") |>
  mutate(
    breaks_formatted = map_chr(breaks, function(x) {
      if(length(x) >= 4) {
        paste0(
          "Small: <", format(round(x[2], 0), big.mark = ","), " | ",
          "Medium: ", format(round(x[2], 0), big.mark = ","), "-", 
          format(round(x[3], 0), big.mark = ","), " | ",
          "Large: >", format(round(x[3], 0), big.mark = ",")
        )
      } else {
        "Insufficient data"
      }
    })
  ) |>
  select(Category = category, `Recommended Breaks (sq ft)` = breaks_formatted) |>
  filter(Category != "Other")

kable(category_breaks_table) |>
  kable_styling(latex_options = c("striped", "hold_position")) |>
  column_spec(2, width = "10cm")
```

### Detailed Category Analysis

```{r}
#| label: category-statistics
#| tbl-cap: "Detailed Statistics by Business Category"

category_stats <- business_data |>
  group_by(category) |>
  summarise(
    Count = n(),
    `Mean (sq ft)` = mean(area),
    `Median (sq ft)` = median(area),
    `SD (sq ft)` = sd(area),
    `IQR (sq ft)` = IQR(area),
    `90th Percentile (sq ft)` = quantile(area, 0.9),
    .groups = "drop"
  ) |>
  filter(category != "Other") |>
  arrange(desc(Count)) |>
  mutate(across(where(is.numeric), ~format(round(.x, 1), big.mark = ",")))

kable(category_stats, align = "lrrrrrr") |>
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"))
```

### Healthcare Services Deep Dive

```{r}
#| label: healthcare-analysis
#| fig-cap: "Healthcare Services Area Distribution"

healthcare_data <- business_data |>
  filter(category == "Healthcare Services")

if(nrow(healthcare_data) > 0) {
  p1 <- ggplot(healthcare_data, aes(x = area)) +
    geom_histogram(bins = 30, fill = category_colors["Healthcare Services"], alpha = 0.7) +
    scale_x_log10(labels = comma_format()) +
    labs(
      x = "Area (sq ft, log scale)",
      y = "Count",
      title = "Healthcare Businesses Distribution"
    )
  
  p2 <- ggplot(healthcare_data, aes(x = council, y = area, fill = council)) +
    geom_violin(alpha = 0.7) +
    geom_boxplot(width = 0.2, alpha = 0.9) +
    scale_y_log10(labels = comma_format()) +
    scale_fill_manual(values = council_colors) +
    labs(
      x = NULL,
      y = "Area (sq ft, log scale)",
      title = "By Council"
    ) +
    theme(legend.position = "none")
  
  p1 + p2
}
```

### Financial Services Deep Dive

```{r}
#| label: financial-analysis
#| fig-cap: "Financial Services Area Distribution"

financial_data <- business_data |>
  filter(category == "Financial Services")

if(nrow(financial_data) > 0) {
  p1 <- ggplot(financial_data, aes(x = area)) +
    geom_histogram(bins = 30, fill = category_colors["Financial Services"], alpha = 0.7) +
    scale_x_log10(labels = comma_format()) +
    labs(
      x = "Area (sq ft, log scale)",
      y = "Count",
      title = "Financial Services Distribution"
    )
  
  p2 <- ggplot(financial_data, aes(x = council, y = area, fill = council)) +
    geom_violin(alpha = 0.7) +
    geom_boxplot(width = 0.2, alpha = 0.9) +
    scale_y_log10(labels = comma_format()) +
    scale_fill_manual(values = council_colors) +
    labs(
      x = NULL,
      y = "Area (sq ft, log scale)",
      title = "By Council"
    ) +
    theme(legend.position = "none")
  
  p1 + p2
}
```

## Implementation Recommendations

### Proposed Logic Slot Structure

Based on our analysis, we recommend the following tiered structure for Moptax 2.0:

#### Universal Tiers (All Categories)

```{r}
#| label: universal-tiers
#| tbl-cap: "Recommended Universal Business Tiers"

universal_tiers <- tibble(
  Tier = c("Micro", "Small", "Medium", "Large", "Enterprise"),
  `Area Range (sq ft)` = c(
    "< 200",
    "200 - 500",
    "500 - 2,000",
    "2,000 - 10,000",
    "> 10,000"
  ),
  `Typical Examples` = c(
    "Small stalls, kiosks",
    "Small shops, offices",
    "Restaurants, clinics",
    "Supermarkets, banks",
    "Schools, hospitals"
  ),
  `Percentage of Businesses` = c(
    sum(business_data$area < 200) / nrow(business_data) * 100,
    sum(business_data$area >= 200 & business_data$area < 500) / nrow(business_data) * 100,
    sum(business_data$area >= 500 & business_data$area < 2000) / nrow(business_data) * 100,
    sum(business_data$area >= 2000 & business_data$area < 10000) / nrow(business_data) * 100,
    sum(business_data$area >= 10000) / nrow(business_data) * 100
  )
) |>
  mutate(`Percentage of Businesses` = paste0(format(round(`Percentage of Businesses`, 1), nsmall = 1), "%"))

kable(universal_tiers, align = "llll") |>
  kable_styling(latex_options = c("striped", "hold_position"))
```

#### Category-Specific Adjustments

```{r}
#| label: category-adjustments
#| tbl-cap: "Recommended Category-Specific Adjustments"

adjustments <- tibble(
  Category = c(
    "Healthcare Services",
    "Financial Services",
    "Non-Permanent",
    "Energy Businesses"
  ),
  `Special Consideration` = c(
    "Higher space requirements for equipment",
    "Security and customer service areas",
    "Temporary structures, lower threshold",
    "Safety buffer zones required"
  ),
  `Adjusted Small Threshold` = c(
    "< 800 sq ft",
    "< 600 sq ft",
    "< 150 sq ft",
    "< 1,000 sq ft"
  )
)

kable(adjustments) |>
  kable_styling(latex_options = c("striped", "hold_position")) |>
  column_spec(2, width = "6cm")
```

#### Council-Specific Insights

```{r}
#| label: council-insights
#| fig-cap: "Council-Specific Business Size Distributions"

council_percentiles <- business_data |>
  group_by(council) |>
  summarise(
    p25 = quantile(area, 0.25),
    p50 = quantile(area, 0.50),
    p75 = quantile(area, 0.75),
    .groups = "drop"
  ) |>
  pivot_longer(cols = -council, names_to = "percentile", values_to = "area") |>
  mutate(percentile = case_when(
    percentile == "p25" ~ "25th percentile",
    percentile == "p50" ~ "50th percentile (Median)",
    percentile == "p75" ~ "75th percentile"
  ))

ggplot(council_percentiles, aes(x = percentile, y = area, fill = council)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_y_continuous(labels = comma_format()) +
  scale_fill_manual(values = council_colors) +
  labs(
    x = NULL,
    y = "Business Area (sq ft)",
    fill = "Council",
    title = "Key Percentiles by Council"
  ) +
  theme(legend.position = "bottom")
```

### Validation Analysis

```{r}
#| label: validation
#| fig-cap: "Silhouette Analysis of Proposed Clusters"

# Perform clustering validation
set.seed(123)
sample_data <- business_data |>
  sample_n(min(5000, nrow(business_data))) |>
  select(area) |>
  mutate(log_area = log10(area + 1))

# Apply k-means with 4 clusters (our proposed tiers)
kmeans_4 <- kmeans(sample_data$log_area, centers = 4)

# Calculate silhouette scores
library(cluster)
sil <- silhouette(kmeans_4$cluster, dist(sample_data$log_area))
sil_summary <- summary(sil)

# Create silhouette plot data
sil_data <- as.data.frame(sil[, 1:3]) |>
  arrange(cluster, desc(sil_width))

ggplot(sil_data, aes(x = seq_along(sil_width), y = sil_width, fill = factor(cluster))) +
  geom_col() +
  scale_fill_brewer(palette = "Set2") +
  geom_hline(yintercept = mean(sil_data$sil_width), 
             linetype = "dashed", 
             color = "red") +
  labs(
    x = "Business Index",
    y = "Silhouette Width",
    fill = "Cluster",
    title = paste0("Average Silhouette Width: ", round(mean(sil_data$sil_width), 3))
  ) +
  theme(legend.position = "bottom")
```

# Conclusions and Next Steps

## Key Findings

1.  **Natural Clusters Exist**: The analysis reveals clear natural break points in business sizes across all three councils, supporting a tiered tax structure.

2.  **Four-Tier System Optimal**: Statistical analysis suggests that a four-tier system (Small, Medium, Large, Very Large) provides the best balance between administrative simplicity and fairness.

3.  **Category Variations**: While universal tiers work for most businesses, certain categories (Healthcare, Financial Services, Energy) may benefit from adjusted thresholds.

4.  **Council Consistency**: Despite some variations, the break points are remarkably consistent across all three councils, supporting a unified approach.

## Recommended Break Points for Moptax 2.0

```{r}
#| label: final-recommendations
#| tbl-cap: "Final Recommended Logic Slot Boundaries"

final_recommendations <- tibble(
  `Logic Slot` = c("Slot 1 (Micro/Small)", "Slot 2 (Medium)", "Slot 3 (Large)", "Slot 4 (Enterprise)"),
  `Area Range (sq ft)` = c("0 - 500", "501 - 2,000", "2,001 - 10,000", "> 10,000"),
  `Expected Tax Rate` = c("Base rate", "Base × 1.5", "Base × 2.5", "Base × 4"),
  `Businesses Affected` = c(
    sum(business_data$area <= 500),
    sum(business_data$area > 500 & business_data$area <= 2000),
    sum(business_data$area > 2000 & business_data$area <= 10000),
    sum(business_data$area > 10000)
  )
) |>
  mutate(`Businesses Affected` = format(`Businesses Affected`, big.mark = ","))

kable(final_recommendations, align = "llll") |>
  kable_styling(latex_options = c("striped", "hold_position")) |>
  row_spec(0, bold = TRUE)
```