---
title: "Delivery Pace Analysis: Freetown, Kenema, and Makeni" 
author: "Lorena Edah"
format: pdf
editor: visual
---

```{=latex}
\begin{figure}[H]
\centering
\includegraphics[width=0.6\textwidth]{cover_page.png}
\end{figure}
```

# **Summary**

#### **Objective**

To determine a realistic and accurate daily delivery pace to set for each city—Freetown, Kenema, and Makeni—based on past delivery data.

#### Key Findings

-   **Freetown** data shows the highest capacity (64.30 deliveries/day in 2025) with gradual long-term growth averaging 2.21% annually, but with some year-to-year fluctuations.
-   **Kenema** data demonstrates significant improvement in average daily deliveries with a 4.49% growth from 2024 to 2025. Kenema's data went from an inconsistent bimodal distribution in 2024 to stable and predictable averaging 40.96 deliveries/day in 2025.
-   **Makeni** maintains steady baseline performance at 31.68 deliveries/day with predictable patterns and minimal variability.

**Operational Patterns Across All 3 Cities:**

-   Peak productivity occurs during mid-day hours (12:00-14:00) across all cities.
-   Mid-week days demonstrate highest delivery volumes.
-   Filtering for delivery accuracy has minimal impact on performance metrics.

## Policy Recommendations

#### 2026 Daily Delivery Targets

Targets are set as ranges, with a minimum to sustain productivity and a maximum to reward high achievers without compromising performance quality.

-   **Makeni**: 33-50 deliveries per day
-   **Kenema**: 43-54 deliveries per day
-   **Freetown**: 69-82 deliveries per day

#### Strategic Considerations

-   **Growth Management**: Kenema and Makeni show capacity for continued expansion, while Freetown is slowly reaching sustainable growth management given its mature operational status.
-   **Target Setting Framework**: Future targets should incorporate field staff consultation and continuous data monitoring.
-   **Fixed Rate Transition**: Consider maintaining fixed targets when average annual growth falls below 2%, operational strain emerges, or distributions stabilize.

# **Methodology**

The datasets used for this analysis were obtained in CSV format. They include: mcc_2025, kcc_2024, kcc_2025, fcc_2021_team_1-3, fcc_2021_team_1-3, fcc_2021_team_4-6, fcc_2022_team_1-3, fcc_2022_team_4-6, fcc_2023_team_1-3, fcc_2023_team_4-6, fcc_2024_internal_staff, fcc_2024_team_1-3,fcc_2024_team_4-6, fcc_2025_rnd1_team_1-3, and fcc_2025_rnd1_team_4-6.

To establish reliable delivery pace benchmarks, the following procedures were applied systematically to each city and year of data:

1.  **Comparing Filtered vs. Unfiltered Data:** To understand how accuracy filtering affected the number of daily deliveries enumerators completed.
    -   For comparison purposes, two subdatasets were created from the initial dataset—an unfiltered version containing all recorded deliveries and a filtered version restricted to accurate deliveries. Deliveries were classified as accurate if they occurred within 80 meters of the assigned property. This threshold served as a proxy for correct and successfully completed deliveries.
    -   The data sets were then grouped by enumerator and date to determine the number of total and accurate deliveries each enumerator completed per day.
    -   To ensure results reflected typical working days, extreme values were removed using the Interquartile Range (IQR) method. This prevented unusually high or low delivery counts from skewing the analysis.
    -   After cleaning, descriptive statistics were calculated, including the mean, median, standard deviation, interquartile range, and benchmarks at the 80th and 90th percentiles. These provided a balanced picture of typical and high-end performance.
    -   Key indicators from the filtered and unfiltered datasets were compared directly. Percentage differences between the two revealed the extent to which accuracy filtering influenced daily delivery averages.
2.  **Temporal Analysis:** To understand **when** enumerators were most productive, deliveries were examined across different timeframes:
    -   **By time of Day:** Deliveries were categorized into three-hour intervals to pinpoint peak working hours.
    -   **By Day of Week:** Average delivery pace was compared across days to detect patterns in weekday versus weekend productivity.
    -   **By Day and Time Combined:** Delivery pace was analyzed jointly by day of week and time of day to identify the most productive periods overall.
3.  **Visualization of Results:** Key findings were summarized through clear visuals, including time-series plots to show trends, histograms with density curves to illustrate distributions, and bar charts to highlight productivity by day and time brackets.

**Note:** For the temporal analysis and visualizations, the **unfiltered dataset** was used. This was an executive decision to present delivery pace in its entirety rather than limiting the view to only accurate deliveries. Since errors are an expected part of the process, including them provides a more realistic basis for understanding overall performance and setting daily pace targets.

**Replication Across Cities and Years:** The full process was applied consistently to each city and each year of available data. This enabled both city-specific insights and cross-city comparisons to identify broader trends and differences.

# Contextual Background

The three cities analyzed in this report are at different stages of implementing the property tax reform, which directly affects the scope and structure of available delivery data.

-   **Makeni** introduced the property tax reform most recently, in 2024–2025. As a result, only one year of delivery data is available for analysis.
-   **Kenema** has two years of delivery data, reflecting its earlier adoption of the reform.
-   **Freetown**, with the longest experience, has five years of delivery data, providing the most comprehensive dataset.

In addition to differences in data length, the underlying data structures vary by city due to the use of different versions of the MOPTAX system:

-   **Makeni’s** data was captured entirely in **MOPTAX 2.0**.
-   **Kenema’s** data spans both **MOPTAX 1.0 and 2.0**, creating a hybrid structure.
-   **Freetown’s** data was recorded exclusively in **MOPTAX 1.0**.

These variations matter because the two MOPTAX versions create datasets that differ in column names, value formats, and certain metrics. Consequently, the analysis required careful adjustments to account for system-specific differences and ensure comparability across cities.

Moreover, Freetown’s 2025 dataset contains additional complexities, as it includes two rounds of delivery. This analysis focuses only on Round 1, which primarily consisted of residential (RDN) deliveries and excluded business licenses. Business license deliveries often take a different amount of time than residential deliveries, thus this data irregularity could distort delivery pace outcomes. As such, results from Freetown 2025 should be interpreted with this context in mind.

Finally, also important to note is that most typical delivery period work days span from Monday or Tuesday through Friday or Saturday, and should almost always fall within standard operational hours of 8:00 a.m. to 5:30 p.m. This has implications for deriving insights from the temporal analysis conducted.

```{r}
#| label: Loading the Necessary Packages
#| include: false

library(tidyverse)
library(janitor)
library(skimr)
library(ggthemes)
library(knitr)
library(glue)

```

```{r}
#| label: Makeni 2025 Analysis
#| include: false

# Loading and cleaning the data
mcc_2025 <- read_csv("mcc_2025.csv") |> 
  clean_names() # standardize column names

# Transforming: unfiltered data (all deliveries regardless of accuracy)
mcc_2025_unfiltered <- mcc_2025 |> 
  mutate(date = as_date(delivered_on)) |> # create a proper date column
  count(date, mobile_user_username, name = "enum_daily_deliv") |> # count deliveries per enumerator per day
  group_by(date) |> # # group all enumerators together by date
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> # average across enumerators per day
  mutate( # Removing outliers using the IQR method
  Q1 = quantile(daily_avg, 0.25),
  Q3 = quantile(daily_avg, 0.75),
  IQR = Q3 - Q1
) |>
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Transforming: filtered data (only accurate deliveries <= 80m) 
mcc_2025_filtered <- mcc_2025 |> 
  filter(distance <= 80) |> # keep only accurate deliveries (<= 80m threshold)
  mutate(date = as_date(delivered_on)) |> 
  count(date, mobile_user_username, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate( 
  Q1 = quantile(daily_avg, 0.25),
  Q3 = quantile(daily_avg, 0.75),
  IQR = Q3 - Q1
) |>
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Summary statistics 
makeni_stats_unfiltered <- mcc_2025_unfiltered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),    # average daily pace
    median = median(daily_avg, na.rm = TRUE),  # middle point (robust to outliers)
    sd     = sd(daily_avg, na.rm = TRUE),      # variation
    iqr    = IQR(daily_avg, na.rm = TRUE),     # spread of typical days
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE), # 80th percentile (ambitious target)
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)  # 90th percentile (high benchmark)
  )
table_text <- capture.output(
  kable(makeni_stats_unfiltered, format = "simple", caption = "Makeni 2025 Unfiltered Data: Average Delivery Pace")
)
writeLines(table_text, "mcc_2025_delivery_pace_unfiltered.txt")
View(makeni_stats_unfiltered)

makeni_stats_filtered <- mcc_2025_filtered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),    
    median = median(daily_avg, na.rm = TRUE),  
    sd     = sd(daily_avg, na.rm = TRUE),      
    iqr    = IQR(daily_avg, na.rm = TRUE),     
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE), 
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)  
  )
table_text <- capture.output(
  kable(makeni_stats_filtered, format = "simple", caption = "Makeni 2025 Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "mcc_2025_delivery_pace_filtered.txt")
View(makeni_stats_filtered)

# Compare unfiltered vs filtered averages 
makeni_comparison_2025 <- tibble(
  unfiltered_mean = mean(mcc_2025_unfiltered$daily_avg, na.rm = TRUE), 
  filtered_mean   = mean(mcc_2025_filtered$daily_avg, na.rm = TRUE) 
) |> 
  mutate(
    diff_percent = (unfiltered_mean - filtered_mean) / unfiltered_mean * 100 # % difference: all deliveries vs. accurate deliveries only
  )
table_text <- capture.output(
  kable(makeni_comparison_2025, format = "simple", caption = "Makeni 2025 Unfiltered vs. Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "mcc_2025_delivery_pace_comparison.txt")
View(makeni_comparison_2025)

# Looking at How Time of Day & Day of Week Affects Delivery Pace
mcc_2025_tod_dow <- mcc_2025 |> 
  mutate(
    date = as_date(delivered_on), # convert to proper date format
    dow = wday(date, label = TRUE, abbr = TRUE), # extract day of week as abbreviated labels (Mon, Tue, etc.)
    hour = hour(delivered_on), # extract hour of day from timestamp
    tod = case_when( # create time-of-day brackets (3-hour periods)
      hour >= 6  & hour < 9  ~ "06–08", # early morning shift
      hour >= 9  & hour < 12 ~ "09–11", # mid-morning shift
      hour >= 12 & hour < 15 ~ "12–14", # early afternoon shift
      hour >= 15 & hour < 18 ~ "15–17", # late afternoon shift
      TRUE ~ "Other (Night/Early Morning)" # unusual hours
    )
  )

# Calculate average delivery pace by time of day
mcc_2025_delivery_pace_tod <- mcc_2025_tod_dow |> 
  count(date, tod, mobile_user_username, name = "enum_deliv") |> # count deliveries per enumerator per time of day (tod)
  group_by(tod) |> # group by day of week
  summarise(
    avg_daily_pace_tod = mean(enum_deliv),
    .groups = "drop" 
  )
table_text <- capture.output(
  kable(mcc_2025_delivery_pace_tod, format = "simple", caption = "Makeni 2025 Time of Day: Average Delivery Pace")
)
writeLines(table_text, "mcc_2025_delivery_pace_tod.txt")
View(mcc_2025_delivery_pace_tod)

# Calculate average delivery pace by day of week
mcc_2025_delivery_pace_dow <- mcc_2025_tod_dow |> 
  count(date, dow, mobile_user_username, name = "enum_deliv") |> # count deliveries per enumerator per day of the week (dow)
  group_by(dow) |> # group by day of week
  summarise(
    avg_daily_pace_dow = mean(enum_deliv),
    .groups = "drop" 
  )
table_text <- capture.output(
  kable(mcc_2025_delivery_pace_dow, format = "simple", caption = "Makeni 2025 Day of Week: Average Delivery Pace")
)
writeLines(table_text, "mcc_2025_delivery_pace_dow.txt")
View(mcc_2025_delivery_pace_dow)


# Calculate average delivery pace by both day of week and time of day
mcc_2025_delivery_pace_tod_dow <- mcc_2025_tod_dow |> 
  count(date, dow, tod, mobile_user_username, name = "enum_deliv") |> 
  group_by(dow, tod) |> 
  summarise(
    avg_daily_pace_tod_dow = mean(enum_deliv), 
    .groups = "drop" 
  )
  
# Visualizations 
# (a) Time-series plot (daily delivery averages over time)
mcc_2025_unfiltered |> 
  ggplot(aes(x = date, y = daily_avg)) + 
  geom_line(color = "blue", linewidth = 1) + 
  geom_point(size = 1.5, color = "black") +  
  labs(                                      
    title = "Makeni 2025: Average Accurate Daily Deliveries", 
    x = "Date",                            
    y = "Average Accurate Daily Deliveries"
  ) 
ggsave("makeni_2025_timeseries.png")  

# (b) Histogram + density curve (distribution shape)
mcc_2025_unfiltered |> 
  ggplot(aes(x = daily_avg)) +            
  geom_histogram(aes(y = after_stat(density)), 
                  bins = 15, fill = "lightgreen", alpha = 0.6, color = "black") + 
  geom_density(color = "darkgreen", linewidth = 1) + 
  labs(                                     
    title = "Distribution of Daily Delivery Pace (Makeni 2025)", 
    x = "Deliveries per Day",              
    y = "Density"                         
  )
ggsave("makeni_2025_distribution.png")  
      
# (c) Delivery pace by day of week and time of day bracket
mcc_2025_delivery_pace_tod_dow |>                
  ggplot(aes(x = dow, y = avg_daily_pace_tod_dow, fill = tod)) + 
  geom_col(position = "dodge") +        
  labs(                                 
    title = "Makeni 2025: Delivery Pace by Day of Week & Time of Day", # Chart title
    x = "Day of the Week",              
    y = "Average Deliveries per Enumerator", 
    fill = "Time of Day (3-hour brackets)" 
  )
ggsave("makeni_2025_pace_dow_tod.png") 

writeLines(strwrap(" 
Summary Characteristics: Filtered vs. Unfiltered Data 

The makeni data analysis showcases impressively consistent performance when comparing both filtered and unfiltered datasets. The mean daily deliveries stand at 31.68 for unfiltered data compared to 31.28 for filtered data, representing only a minimal 1.24% difference.The median performance figures of 33.09 deliveries for unfiltered data versus 32.60 for filtered data further reinforce this consistency, while the standard deviation of approximately 14.6 to 14.8 deliveries indicates moderate but acceptable variability in daily performance. When examining high-performance thresholds, the data reveals that the 80th percentile reaches 45 to 46 deliveries per day, while the 90th percentile extends to 48 or more deliveries, establishing clear benchmarks for exceptional operational days.


Temporal Performance Patterns

Day of Week Analysis

Peak Performance Days represent the operational sweet spot where multiple factors align to create optimal delivery conditions. Tuesday emerges as the standout performer with 48.66 average daily deliveries, establishing itself as the highest-performing day of the week. Monday follows closely with 43.57 average daily deliveries, followed by Friday which has a strong performance of 36.30 average daily deliveries. These three days collectively demonstrate the operation's capacity to handle increased volume while maintaining service quality.

Moderate Performance Days include Wednesday with an average daily delivery pace of 29.2, followed closely by Thursday with 27.44 average daily deliveries, and finally Saturday with 23.41 deliveries. These days likely represent periods where demand patterns, staffing levels, or logistical factors create more standard operational conditions.

Sundays are the lowest performance days with an average of  13.80 daily deliveries. This dramatic reduction is to be expected as enumerators should not be working on Sundays.

Time of Day Performance

The 12-14 hour period stands out as the peak performance window, achieving 18.39 deliveries and representing the most productive time slot of the operational day. The morning period from 09-11 hours follows with a strong performance of  14.94 average deliveries. The afternoon window from 15-17 hours follows with an average of 12.89 deliveries.

Lower Activity Periods include early morning periods from 06-08 hours with an average of 5.91 deliveries followed by rhe night and early morning periods outside standard hours with an average of 4.97 deliveries all of which is to be expected as enumerators should not be working these hours.

Visualizations

Distribution Analysis
This distribution pattern displays multiple performance clusters with clear peaks occurring around 10, 30, 40, and 50 deliveries per day, indicating that average delivery pace evolved in distinct performance modes rather than following a simple linear progression as the distribution spans from approximately 5 deliveries up to 50 or more deliveries per day. 

Temporal Trend Analysis

The time series analysis reveals significant day-to-day variability that tells a story of dynamic average delivery pace. The performance range extends dramatically from a low of 7 deliveries to a peak of 51 deliveries per day, showcasing the full spectrum of operational capability. Notable performance peaks occur on April 14-15, where the operation achieves over 50 deliveries per day. The data also shows significant performance drops during mid-period and at month-end. However, there are consistent recovery patterns that follow these drops.", width = 80))


```

```{r}
#| label: Kenema 2025 Analysis
#| include: false

# Loading and cleaning the data
kcc_2025 <- read_csv("kcc_2025.csv") |> 
  clean_names() 
    
# Transforming: unfiltered data (all deliveries, regardless of accuracy)
kcc_2025_unfiltered <- kcc_2025 |> 
  mutate(date = as_date(mdy_hm(delivered_on))) |>
  count(date, user_name, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate( 
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |>
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )
    
# Transforming: filtered data (only accurate deliveries <= 80m) 
kcc_2025_filtered <- kcc_2025 |> 
  filter(distance <= 80) |> 
  mutate(date = as_date(mdy_hm(delivered_on))) |> 
  count(date, user_name, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate( 
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |>
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )
    
# Summary statistics 
kenema_stats_unfiltered <- kcc_2025_unfiltered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),    
    median = median(daily_avg, na.rm = TRUE),  
    sd     = sd(daily_avg, na.rm = TRUE),      
    iqr    = IQR(daily_avg, na.rm = TRUE),     
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE) 
  )
table_text <- capture.output(
  kable(kenema_stats_unfiltered, format = "simple", caption = "Kenema 2025 Unfiltered Data: Average Delivery Pace")
)
writeLines(table_text, "kcc_2025_delivery_pace_unfiltered.txt")
View(kenema_stats_unfiltered)

    
kenema_stats_filtered <- kcc_2025_filtered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),    
    median = median(daily_avg, na.rm = TRUE),  
    sd     = sd(daily_avg, na.rm = TRUE),      
    iqr    = IQR(daily_avg, na.rm = TRUE),     
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE), 
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)  
  )
table_text <- capture.output(
  kable(kenema_stats_filtered, format = "simple", caption = "Kenema 2025 Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "kcc_2025_delivery_pace_filtered.txt")
View(kenema_stats_filtered)

    
# Compare unfiltered vs filtered averages
kenema_comparison_2025 <- tibble(
  unfiltered_mean = mean(kcc_2025_unfiltered$daily_avg, na.rm = TRUE),
  filtered_mean   = mean(kcc_2025_filtered$daily_avg, na.rm = TRUE)
) |> 
  mutate(
    diff_percent = (unfiltered_mean - filtered_mean) / unfiltered_mean * 100
  )
table_text <- capture.output(
  kable(kenema_comparison_2025, format = "simple", caption = "Kenema 2025 Unfiltered vs. Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "kcc_2025_delivery_pace_comparison.txt")
View(kenema_comparison_2025)

    
# Looking at How Time of Day & Day of Week Affects Delivery Pace
kcc_2025_tod_dow <- kcc_2025 |> 
  mutate(
    datetime = mdy_hm(delivered_on),          
    date = as_date(datetime),                 
    dow = wday(date, label = TRUE, abbr = TRUE), 
    hour = hour(datetime),                     
    tod = case_when(                          
      hour >= 6  & hour < 9  ~ "06–08",
      hour >= 9  & hour < 12 ~ "09–11",
      hour >= 12 & hour < 15 ~ "12–14",
      hour >= 15 & hour < 18 ~ "15–17",
      TRUE ~ "Other (Night/Early Morning)"
    )
  )
    
# Calculate average delivery pace by time of day
kcc_2025_delivery_pace_tod <- kcc_2025_tod_dow |> 
  count(date, tod, user_name, name = "enum_deliv") |> 
  group_by(tod) |> 
  summarise(
    avg_daily_pace_tod = mean(enum_deliv), 
    .groups = "drop"
  )
table_text <- capture.output(
  kable(kcc_2025_delivery_pace_tod, format = "simple", caption = "Kenema 2025 Time of Day: Average Delivery Pace")
)
writeLines(table_text, "kcc_2025_delivery_pace_tod.txt")
View(kcc_2025_delivery_pace_tod)

    
# Calculate average delivery pace by day of week
kcc_2025_delivery_pace_dow <- kcc_2025_tod_dow |> 
  count(date, dow, user_name, name = "enum_deliv") |> 
  group_by(dow) |> 
  summarise(
    avg_daily_pace_dow = mean(enum_deliv), 
    .groups = "drop"
  )
table_text <- capture.output(
  kable(kcc_2025_delivery_pace_dow, format = "simple", caption = "Kenema 2025 Day of Week: Average Delivery Pace")
)
writeLines(table_text, "kcc_2025_delivery_pace_dow.txt")
View(kcc_2025_delivery_pace_dow)
    
# Calculate average delivery pace by both day of week and time of day
kcc_2025_delivery_pace_tod_dow <- kcc_2025_tod_dow |> 
  count(date, dow, tod, user_name, name = "enum_deliv") |> 
  group_by(dow, tod) |> 
  summarise(
    avg_daily_pace_tod_dow = mean(enum_deliv), 
    .groups = "drop"
  )

# Visualizations 
# (a) Time-series plot (performance over time)
kcc_2025_unfiltered |> 
  ggplot(aes(x = date, y = daily_avg)) +
  geom_line(color = "blue", linewidth = 1) + 
  geom_point(size = 1.5, color = "black") +       
  labs(
    title = "Kenema 2025: Average Accurate Daily Deliveries",
    x = "Date",
    y = "Average Accurate Daily Deliveries"
  ) 
ggsave("kenema_2025_timeseries.png")
    
# (b) Histogram + density curve (distribution shape)
kcc_2025_unfiltered |> 
  ggplot(aes(x = daily_avg)) +
  geom_histogram(aes(y = after_stat(density)), 
                 bins = 15, fill = "lightcoral", alpha = 0.6, color = "black") +
  geom_density(color = "darkred", linewidth = 1) +
  labs(
    title = "Distribution of Daily Delivery Pace (Kenema 2025)",
    x = "Deliveries per Day",
    y = "Density"
  ) 
ggsave("kenema_2025_distribution.png")
    
# (c) Delivery pace by day of week and time of day bracket
kcc_2025_delivery_pace_tod_dow |> 
  ggplot(aes(x = dow, y = avg_daily_pace_tod_dow, fill = tod)) +
  geom_col(position = "dodge") +
  labs(
    title = "Kenema 2025: Delivery Pace by Day of Week & Time of Day",
    x = "Day of the Week",
    y = "Average Deliveries per Enumerator",
    fill = "Time of Day (3-hour brackets)"
  ) 
ggsave("kenema_2025_pace_dow_tod.png")

writeLines(strwrap(" 
Summary Characteristics: Filtered vs. Unfiltered Data

The Kenema data analysis showcases impressively consistent performance when comparing both filtered and unfiltered datasets. The mean daily deliveries stand at 40.96 for unfiltered data compared to 40.15 for filtered data, representing only a minimal 1.98% difference. The median performance figures of 41.1 deliveries for unfiltered data versus 40.43 for filtered data further reinforce this consistency, while the standard deviation of approximately 8.77 to 9.17 deliveries indicates moderate but acceptable variability in daily performance. When examining high-performance thresholds, the data reveals that the 80th percentile reaches 48 to 49 deliveries per day, while the 90th percentile extends to 50 or more deliveries, establishing clear benchmarks for exceptional operational days.

Temporal Performance Patterns

Day of Week Analysis
Peak Performance Days represent the operational sweet spot where multiple factors align to create optimal delivery conditions. Sunday emerges as the standout performer with 49.50 average daily deliveries, establishing itself as the highest-performing day of the week. Wednesday follows closely with 46.36 average daily deliveries, followed by Thursday which has a strong performance of 45.70 average daily deliveries. These three days collectively demonstrate the operation's capacity to handle increased volume while maintaining service quality.

Moderate Performance Days include Tuesday with an average daily delivery pace of 43.01, followed closely by Monday with 38.80 average daily deliveries, and finally Friday with 36.34 deliveries. 

Saturday is the lowest performance day with an average of 33.52 daily deliveries. This reduction represents the lowest operational performance point in the weekly cycle.

Time of Day Performance

The 12-14 hour period stands out as the peak performance window, achieving 20.55 deliveries and representing the most productive time slot of the operational day. The morning period from 09-11 hours follows with a strong performance of 18.00 average deliveries. The afternoon window from 15-17 hours follows with an average of 14.71 deliveries.

Lower Activity Periods include early morning periods from 06-08 hours with an average of 5.97 deliveries followed by the night and early morning periods outside standard hours with an average of 6.72 deliveries all of which is to be expected as enumerators should not be working these hours.

Visualizations

Distribution Analysis

This distribution pattern displays multiple performance clusters with clear peaks occurring around 20, 35, 40, and 50 deliveries per day, indicating that average delivery pace evolved in distinct performance modes rather than following a simple linear progression as the distribution spans from approximately 15 deliveries up to 60 or more deliveries per day.

Temporal Trend Analysis

The time series analysis reveals significant day-to-day variability that tells a story of dynamic average delivery pace. The performance range extends dramatically from a low of 18 deliveries to a peak of 60 deliveries per day, showcasing the full spectrum of operational capability. Notable performance peaks occur throughout the operational period, where the operation achieves over 50 deliveries per day. The data also shows significant performance drops during various periods. However, there are consistent recovery patterns that follow these drops.", width = 80))

```

```{r}
#| label: Kenema 2024 Analysis
#| include: false

# Loading & cleaning the data
kcc_2024 <- read_csv("kcc_2024.csv")

# Split property_type into multiple rows (each counts as a delivery)
kcc_2024 <- kcc_2024 |> 
  separate_rows(property_type, sep = ",\\s*")

# Transforming: unfiltered data (all deliveries)
kcc_2024_unfiltered <- kcc_2024 |> 
  filter(!is.na(time_delivery_date)) |> # getting rid of missing delivery date data
  mutate(
    datetime = dmy_hms(time_delivery_date),
    date = as_date(datetime)
  ) |> 
  count(date, enteredby, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate(
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |> 
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Transforming: filtered data (accurate deliveries ≤ 0.08 km)
kcc_2024_filtered <- kcc_2024 |> 
  filter(time_gps_distance_original <= 0.08) |> 
  filter(!is.na(time_delivery_date)) |>
  mutate(
    datetime = dmy_hms(time_delivery_date),
    date = as_date(datetime)
  ) |> 
  count(date, enteredby, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate(
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |> 
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Summary statistics
kenema_stats_unfiltered <- kcc_2024_unfiltered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),
    median = median(daily_avg, na.rm = TRUE),
    sd     = sd(daily_avg, na.rm = TRUE),
    iqr    = IQR(daily_avg, na.rm = TRUE),
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)
  )
table_text <- capture.output(
  kable(kenema_stats_unfiltered, format = "simple", caption = "Kenema 2024 Unfiltered Data: Average Delivery Pace")
)
writeLines(table_text, "kcc_2024_delivery_pace_unfiltered.txt")
View(kenema_stats_unfiltered)

kenema_stats_filtered <- kcc_2024_filtered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),
    median = median(daily_avg, na.rm = TRUE),
    sd     = sd(daily_avg, na.rm = TRUE),
    iqr    = IQR(daily_avg, na.rm = TRUE),
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)
  )
table_text <- capture.output(
  kable(kenema_stats_filtered, format = "simple", caption = "Kenema 2024 Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "kcc_2024_delivery_pace_filtered.txt")
View(kenema_stats_filtered)

# Comparison
kenema_comparison_2024 <- tibble(
  unfiltered_mean = mean(kcc_2024_unfiltered$daily_avg, na.rm = TRUE),
  filtered_mean   = mean(kcc_2024_filtered$daily_avg, na.rm = TRUE)
) |> 
  mutate(
    diff_percent = (unfiltered_mean - filtered_mean) / unfiltered_mean * 100
  )
table_text <- capture.output(
  kable(kenema_comparison_2024, format = "simple", caption = "Kenema 2024 Unfiltered vs. Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "kcc_2024_delivery_pace_comparison.txt")
View(kenema_comparison_2024)

# Time of day & day of week
kcc_2024_tod_dow <- kcc_2024 |> 
  filter(!is.na(time_delivery_date)) |>
  mutate(
    datetime = dmy_hms(time_delivery_date),
    date = as_date(datetime),
    dow = wday(date, label = TRUE, abbr = TRUE),
    hour = hour(datetime),
    tod = case_when(
      hour >= 6  & hour < 9  ~ "06–08",
      hour >= 9  & hour < 12 ~ "09–11",
      hour >= 12 & hour < 15 ~ "12–14",
      hour >= 15 & hour < 18 ~ "15–17",
      TRUE ~ "Other (Night/Early Morning)"
    )
  )

# Delivery pace by time of day, day of week
kcc_2024_delivery_pace_tod <- kcc_2024_tod_dow |> 
  count(date, tod, enteredby, name = "enum_deliv") |> 
  group_by(tod) |> 
  summarise(avg_daily_pace_tod = mean(enum_deliv), .groups = "drop")
table_text <- capture.output(
  kable(kcc_2024_delivery_pace_tod, format = "simple", caption = "Kenema 2024 Time of Day: Average Delivery Pace")
)
writeLines(table_text, "kcc_2024_delivery_pace_tod.txt")
View(kcc_2024_delivery_pace_tod)

kcc_2024_delivery_pace_dow <- kcc_2024_tod_dow |> 
  count(date, dow, enteredby, name = "enum_deliv") |> 
  group_by(dow) |> 
  summarise(avg_daily_pace_dow = mean(enum_deliv), .groups = "drop")
table_text <- capture.output(
  kable(kcc_2024_delivery_pace_dow, format = "simple", caption = "Kenema 2024 Day of Week: Average Delivery Pace")
)
writeLines(table_text, "kcc_2024_delivery_pace_dow.txt")
View(kcc_2024_delivery_pace_dow)

kcc_2024_delivery_pace_tod_dow <- kcc_2024_tod_dow |> 
  count(date, dow, tod, enteredby, name = "enum_deliv") |> 
  group_by(dow, tod) |> 
  summarise(avg_daily_pace_tod_dow = mean(enum_deliv), .groups = "drop")

# Visualizations
# (a) Time-series plot
kcc_2024_unfiltered |> 
  ggplot(aes(x = date, y = daily_avg)) +
  geom_line(color = "blue", linewidth = 1) +
  geom_point(size = 1.5, color = "black") +
  labs(
    title = "Kenema 2024: Average Accurate Daily Deliveries", 
    x = "Date", 
    y = "Average Accurate Daily Deliveries"
  ) 
ggsave("kenema_2024_timeseries.png")

# (b) Histogram + density curve
kcc_2024_unfiltered |> 
  ggplot(aes(x = daily_avg)) +
  geom_histogram(aes(y = after_stat(density)), 
                 bins = 15, fill = "lightcoral", alpha = 0.6, color = "black") +
  geom_density(color = "darkred", linewidth = 1) +
  labs(
    title = "Distribution of Daily Delivery Pace (Kenema 2024)", 
    x = "Deliveries per Day", 
    y = "Density"
  ) 
ggsave("kenema_2024_distribution.png")

# (c) Delivery pace by day of week & time of day
kcc_2024_delivery_pace_tod_dow |> 
  ggplot(aes(x = dow, y = avg_daily_pace_tod_dow, fill = tod)) +
  geom_col(position = "dodge") +
  labs(
    title = "Kenema 2024: Delivery Pace by Day of Week & Time of Day",
    x = "Day of the Week", 
    y = "Average Deliveries per Enumerator", 
    fill = "Time of Day (3-hour brackets)"
  ) 
ggsave("kenema_2024_pace_dow_tod.png")

writeLines(strwrap(" 
Summary Characteristics: Filtered vs. Unfiltered Data

The Kenema 2024 data analysis showcases moderately consistent performance when comparing both filtered and unfiltered datasets. The mean daily deliveries stand at 39.20 for unfiltered data compared to 38.83 for filtered data, representing only a minimal 0.94% difference. The median performance figures of 46.07 deliveries for unfiltered data versus 45.69 for filtered data further reinforce this consistency, while the standard deviation of approximately 21.49 to 21.65 deliveries indicates high variability in daily performance. When examining high-performance thresholds, the data reveals that the 80th percentile reaches 56 to 57 deliveries per day, while the 90th percentile extends to 63 or more deliveries, establishing clear benchmarks for exceptional operational days.

Temporal Performance Patterns

Day of Week Analysis

Peak Performance Days represent the operational sweet spot where multiple factors align to create optimal delivery conditions. Thursday emerges as the standout performer with 61.79 average daily deliveries, establishing itself as the highest-performing day of the week. Monday follows closely with 61.43 average daily deliveries, followed by Friday which has a strong performance of 43.66 average daily deliveries. These three days collectively demonstrate the operation's capacity to handle increased volume while maintaining service quality.

Moderate Performance Days include Tuesday with an average daily delivery pace of 37.73, followed closely by Saturday with 21.74 average daily deliveries. 

Wednesday is the lowest performance day with an average of 21.03 daily deliveries. This dramatic reduction represents the lowest operational performance point in the weekly cycle.

Time of Day Performance

The 12-14 hour period stands out as the peak performance window, achieving 21.24 deliveries and representing the most productive time slot of the operational day. The morning period from 09-11 hours follows with a strong performance of 20.87 average deliveries. The afternoon window from 15-17 hours follows with an average of 15.02 deliveries.

Lower Activity Periods include early morning periods from 06-08 hours with an average of 4.87 deliveries followed by the night and early morning periods outside standard hours with an average of 4.73 deliveries all of which is to be expected as enumerators should not be working these hours.

Visualizations

Distribution Analysis

This distribution pattern displays multiple performance clusters with clear peaks occurring around 20 and 45-50 deliveries per day, indicating that average delivery pace evolved in distinct performance modes rather than following a simple linear progression as the distribution spans from approximately 0 deliveries up to 80 or more deliveries per day. The bimodal distribution suggests operational inconsistency with periods of both low and high performance capability.

Temporal Trend Analysis

The time series analysis reveals extreme day-to-day variability that tells a story of volatile average delivery pace. The performance range extends dramatically from a low of 0 deliveries to a peak of 70 deliveries per day, showcasing the full spectrum of operational capability and limitation. Notable performance peaks occur sporadically throughout the operational period, where the operation achieves over 60 deliveries per day. The data also shows significant performance drops with some days recording near-zero deliveries. However, there are inconsistent recovery patterns that follow these drops.", width = 80))


```

```{r}
#| label: Freetown 2025 Analysis
#| include: false

# Loading and cleaning the data: merging the 2 datasets for FCC 2025 Round 1
fcc_2025 <- bind_rows(
  read_csv("fcc_2025_rnd1_team_1-3.csv"),
  read_csv("fcc_2025_rnd1_team_4-6.csv")
) |> 
  clean_names()

# Split property_type into multiple rows (each counts as a delivery)
fcc_2025 <- fcc_2025 |> 
  separate_rows(property_type, sep = ",\\s*")

# Transforming: unfiltered data (all deliveries)
fcc_2025_unfiltered <- fcc_2025 |> 
  filter(!is.na(time_delivery_date)) |> # getting rid of missing delivery date data
  mutate(
    datetime = dmy_hms(time_delivery_date),
    date = as_date(datetime)
  ) |> 
  count(date, enteredby, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate(
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |> 
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Transforming: filtered data (accurate deliveries ≤ 0.08 km)
fcc_2025_filtered <- fcc_2025 |> 
  filter(time_gps_distance_original <= 0.08) |> 
  filter(!is.na(time_delivery_date)) |>
  mutate(
    datetime = dmy_hms(time_delivery_date),
    date = as_date(datetime)
  ) |> 
  count(date, enteredby, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate(
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |> 
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Summary statistics
freetown_stats_unfiltered <- fcc_2025_unfiltered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),
    median = median(daily_avg, na.rm = TRUE),
    sd     = sd(daily_avg, na.rm = TRUE),
    iqr    = IQR(daily_avg, na.rm = TRUE),
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)
  )
table_text <- capture.output(
  kable(freetown_stats_unfiltered, format = "simple", caption = "Freetown 2025 Unfiltered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2025_delivery_pace_unfiltered.txt")
View(freetown_stats_unfiltered)


freetown_stats_filtered <- fcc_2025_filtered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),
    median = median(daily_avg, na.rm = TRUE),
    sd     = sd(daily_avg, na.rm = TRUE),
    iqr    = IQR(daily_avg, na.rm = TRUE),
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)
  )
table_text <- capture.output(
  kable(freetown_stats_filtered, format = "simple", caption = "Freetown 2025 Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2025_delivery_pace_filtered.txt")
View(freetown_stats_filtered)


# Comparison
freetown_comparison_2025 <- tibble(
  unfiltered_mean = mean(fcc_2025_unfiltered$daily_avg, na.rm = TRUE),
  filtered_mean   = mean(fcc_2025_filtered$daily_avg, na.rm = TRUE)
) |> 
  mutate(
    diff_percent = (unfiltered_mean - filtered_mean) / unfiltered_mean * 100
  )
table_text <- capture.output(
  kable(freetown_comparison_2025, format = "simple", caption = "Freetown 2025 Unfiltered vs. Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2025_delivery_pace_comparison.txt")
View(freetown_comparison_2025)


# Time of day & day of week
fcc_2025_tod_dow <- fcc_2025 |> 
  filter(!is.na(time_delivery_date)) |>
  mutate(
    datetime = dmy_hms(time_delivery_date),
    date = as_date(datetime),
    dow = wday(date, label = TRUE, abbr = TRUE),
    hour = hour(datetime),
    tod = case_when(
      hour >= 6  & hour < 9  ~ "06–08",
      hour >= 9  & hour < 12 ~ "09–11",
      hour >= 12 & hour < 15 ~ "12–14",
      hour >= 15 & hour < 18 ~ "15–17",
      TRUE ~ "Other (Night/Early Morning)"
    )
  )

# Delivery pace by time of day, day of week
fcc_2025_delivery_pace_tod <- fcc_2025_tod_dow |> 
  count(date, tod, enteredby, name = "enum_deliv") |> 
  group_by(tod) |> 
  summarise(avg_daily_pace_tod = mean(enum_deliv), .groups = "drop")
table_text <- capture.output(
  kable(fcc_2025_delivery_pace_tod, format = "simple", caption = "Freetown 2025 Time of Day: Average Delivery Pace")
)
writeLines(table_text, "fcc_2025_delivery_pace_tod.txt")
View(fcc_2025_delivery_pace_tod)

fcc_2025_delivery_pace_dow <- fcc_2025_tod_dow |> 
  count(date, dow, enteredby, name = "enum_deliv") |> 
  group_by(dow) |> 
  summarise(avg_daily_pace_dow = mean(enum_deliv), .groups = "drop")
table_text <- capture.output(
  kable(fcc_2025_delivery_pace_dow, format = "simple", caption = "Freetown 2025 Day of Week: Average Delivery Pace")
)
writeLines(table_text, "fcc_2025_delivery_pace_dow.txt")
View(fcc_2025_delivery_pace_dow)

fcc_2025_delivery_pace_tod_dow <- fcc_2025_tod_dow |> 
  count(date, dow, tod, enteredby, name = "enum_deliv") |> 
  group_by(dow, tod) |> 
  summarise(avg_daily_pace_tod_dow = mean(enum_deliv), .groups = "drop")

# Visualizations
# (a) Time-series plot
fcc_2025_unfiltered |> 
  ggplot(aes(x = date, y = daily_avg)) +
  geom_line(color = "blue", linewidth = 1) +
  geom_point(size = 1.5, color = "black") +
  labs(
    title = "Freetown 2025: Average Accurate Daily Deliveries", 
    x = "Date", 
    y = "Average Accurate Daily Deliveries"
  ) 
ggsave("freetown_2025_timeseries.png")

# (b) Histogram + density curve
fcc_2025_unfiltered |> 
  ggplot(aes(x = daily_avg)) +
  geom_histogram(aes(y = after_stat(density)), 
                 bins = 15, fill = "lightblue", alpha = 0.6, color = "black") +
  geom_density(color = "darkblue", linewidth = 1) +
  labs(
    title = "Distribution of Daily Delivery Pace (Freetown 2025)", 
    x = "Deliveries per Day", 
    y = "Density"
  ) 
ggsave("freetown_2025_distribution.png")

# (c) Delivery pace by day of week & time of day
fcc_2025_delivery_pace_tod_dow |> 
  ggplot(aes(x = dow, y = avg_daily_pace_tod_dow, fill = tod)) +
  geom_col(position = "dodge") +
  labs(
    title = "Freetown 2025: Delivery Pace by Day of Week & Time of Day",
    x = "Day of the Week", 
    y = "Average Deliveries per Enumerator", 
    fill = "Time of Day (3-hour brackets)"
  ) 
ggsave("freetown_2025_pace_dow_tod.png")

writeLines(strwrap(" 
Summary Characteristics: Filtered vs. Unfiltered Data

The Freetown 2025 data analysis showcases moderately consistent performance when comparing both filtered and unfiltered datasets. The mean daily deliveries stand at 64.30 for unfiltered data compared to 59.74 for filtered data, representing a 7.11% difference. The median performance figures of 66.66 deliveries for unfiltered data versus 61.37 for filtered data further reinforce this pattern, while the standard deviation of approximately 9.66 to 10.05 deliveries indicates moderate but acceptable variability in daily performance. When examining high-performance thresholds, the data reveals that the 80th percentile reaches 67 to 72 deliveries per day, while the 90th percentile extends to 70 or more deliveries, establishing clear benchmarks for exceptional operational days.

Temporal Performance Patterns

Day of Week Analysis

Saturday emerges as the standout performer with 69.69 average daily deliveries, establishing itself as the highest-performing day of the week. Friday follows closely with 68.32 average daily deliveries, followed by Thursday which has a strong performance of 67.73 average daily deliveries. These three days collectively demonstrate the operation's capacity to handle increased volume while maintaining service quality.

Moderate Performance Days include Wednesday with an average daily delivery pace of 65.49, followed closely by Monday with 65.17 average daily deliveries, and finally Tuesday with 64.66 deliveries.

Sunday is the lowest performance day with an average of 36.33 daily deliveries. This dramatic reduction is to be expected as enumerators should not be working on Sundays.

Time of Day Performance

The 12-14 hour period stands out as the peak performance window, achieving 30.88 deliveries and representing the most productive time slot of the operational day. The morning period from 09-11 hours follows with a strong performance of 26.97 average deliveries. The afternoon window from 15-17 hours follows with an average of 23.85 deliveries.

Lower Activity Periods include the night and early morning periods outside standard hours with an average of 13.94 deliveries followed by early morning periods from 06-08 hours with an average of 8.67 deliveries all of which is to be expected as enumerators should not be working these hours.

Visualizations

Distribution Analysis

This distribution pattern displays multiple performance clusters with clear peaks occurring around 55, 60, and 70 deliveries per day, indicating that average delivery pace evolved in distinct performance modes rather than following a simple linear progression as the distribution spans from approximately 40 deliveries up to 90 or more deliveries per day.

Temporal Trend Analysis

The time series analysis reveals significant day-to-day variability that tells a story of dynamic average delivery pace. The performance range extends dramatically from a low of 40 deliveries to a peak of 90 deliveries per day, showcasing the full spectrum of operational capability. Notable performance peaks occur throughout the operational period, where the operation achieves over 80 deliveries per day. The data also shows significant performance drops during various periods. However, there are consistent recovery patterns that follow these drops.", width = 80))

```

```{r}
#| label: Freetown 2024 Analysis
#| include: false

# Loading and cleaning the data: merging the 2 datasets for FCC 2024
fcc_2024 <- bind_rows(
  read_csv("fcc_2024_team_1-3.csv"),
  read_csv("fcc_2024_team_4-6.csv")
) |> 
  clean_names()

# Check for date parsing issues
bad_dates <- fcc_2024 |>  
  mutate(datetime = dmy_hms(time_delivery_date)) |> 
  filter(is.na(datetime) & !is.na(time_delivery_date)) |> 
  distinct(time_delivery_date)

# Clean up the time_delivery_date column to ensure formatting consistency
fcc_2024 <- fcc_2024 |> 
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    )
  )

# Split property_type column values into multiple rows (each counts as a delivery)
fcc_2024 <- fcc_2024 |> 
  separate_rows(property_type, sep = ",\\s*") 

# Transforming: unfiltered data (all deliveries)
fcc_2024_unfiltered <- fcc_2024 |> 
  filter(!is.na(time_delivery_date)) |> # getting rid of missing delivery date data
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    ),
    date = as_date(datetime)
  )|> 
  count(date, enteredby, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate(
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |> 
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Transforming: filtered data (accurate deliveries ≤ 0.08 km)
fcc_2024_filtered <- fcc_2024 |> 
  filter(time_gps_distance_original <= 0.08) |> 
  filter(!is.na(time_delivery_date)) |>
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    ),
    date = as_date(datetime)
  ) |> 
  count(date, enteredby, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate(
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |> 
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Summary statistics
freetown_stats_unfiltered <- fcc_2024_unfiltered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),
    median = median(daily_avg, na.rm = TRUE),
    sd     = sd(daily_avg, na.rm = TRUE),
    iqr    = IQR(daily_avg, na.rm = TRUE),
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)
  )
table_text <- capture.output(
  kable(freetown_stats_unfiltered, format = "simple", caption = "Freetown 2024 Unfiltered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2024_delivery_pace_unfiltered.txt")
View(freetown_stats_unfiltered)

freetown_stats_filtered <- fcc_2024_filtered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),
    median = median(daily_avg, na.rm = TRUE),
    sd     = sd(daily_avg, na.rm = TRUE),
    iqr    = IQR(daily_avg, na.rm = TRUE),
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)
  )
table_text <- capture.output(
  kable(freetown_stats_filtered, format = "simple", caption = "Kenema 2024 Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2024_delivery_pace_filtered.txt")
View(freetown_stats_filtered)

# Comparison
freetown_comparison_2024 <- tibble(
  unfiltered_mean = mean(fcc_2024_unfiltered$daily_avg, na.rm = TRUE),
  filtered_mean   = mean(fcc_2024_filtered$daily_avg, na.rm = TRUE)
) |> 
  mutate(
    diff_percent = (unfiltered_mean - filtered_mean) / unfiltered_mean * 100
  )
table_text <- capture.output(
  kable(freetown_comparison_2024, format = "simple", caption = "Kenema 2024 Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2024_delivery_pace_comparison.txt")
View(freetown_comparison_2024)

# Time of day & day of week
fcc_2024_tod_dow <- fcc_2024 |> 
  filter(!is.na(time_delivery_date)) |>
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    ),
    date = as_date(datetime),
    dow = wday(date, label = TRUE, abbr = TRUE),
    hour = hour(datetime),
    tod = case_when(
      hour >= 6  & hour < 9  ~ "06–08",
      hour >= 9  & hour < 12 ~ "09–11",
      hour >= 12 & hour < 15 ~ "12–14",
      hour >= 15 & hour < 18 ~ "15–17",
      TRUE ~ "Other (Night/Early Morning)"
    )
  )

# Delivery pace by time of day, day of week
fcc_2024_delivery_pace_tod <- fcc_2024_tod_dow |> 
  count(date, tod, enteredby, name = "enum_deliv") |> 
  group_by(tod) |> 
  summarise(avg_daily_pace_tod = mean(enum_deliv), .groups = "drop")
table_text <- capture.output(
  kable(fcc_2024_delivery_pace_tod, format = "simple", caption = "Freetown 2024 Time of Day: Average Delivery Pace")
)
writeLines(table_text, "fcc_2024_delivery_pace_tod.txt")
View(fcc_2024_delivery_pace_tod)

fcc_2024_delivery_pace_dow <- fcc_2024_tod_dow |> 
  count(date, dow, enteredby, name = "enum_deliv") |> 
  group_by(dow) |> 
  summarise(avg_daily_pace_dow = mean(enum_deliv), .groups = "drop")
table_text <- capture.output(
  kable(fcc_2024_delivery_pace_dow, format = "simple", caption = "Freetown 2024 Day of Week: Average Delivery Pace")
)
writeLines(table_text, "fcc_2024_delivery_pace_dow.txt")
View(fcc_2024_delivery_pace_dow)

fcc_2024_delivery_pace_tod_dow <- fcc_2024_tod_dow |> 
  count(date, dow, tod, enteredby, name = "enum_deliv") |> 
  group_by(dow, tod) |> 
  summarise(avg_daily_pace_tod_dow = mean(enum_deliv), .groups = "drop")

# Visualizations
# (a) Time-series plot
fcc_2024_unfiltered |> 
  ggplot(aes(x = date, y = daily_avg)) +
  geom_line(color = "blue", linewidth = 1) +
  geom_point(size = 1.5, color = "black") +
  labs(
    title = "Freetown 2024: Average Accurate Daily Deliveries", 
    x = "Date", 
    y = "Average Accurate Daily Deliveries"
  ) 
ggsave("freetown_2024_timeseries.png")

# (b) Histogram + density curve
fcc_2024_unfiltered |> 
  ggplot(aes(x = daily_avg)) +
  geom_histogram(aes(y = after_stat(density)), 
                 bins = 15, fill = "lightblue", alpha = 0.6, color = "black") +
  geom_density(color = "darkblue", linewidth = 1) +
  labs(
    title = "Distribution of Daily Delivery Pace (Freetown 2024)", 
    x = "Deliveries per Day", 
    y = "Density"
  ) 
ggsave("freetown_2024_distribution.png")

# (c) Delivery pace by day of week & time of day
fcc_2024_delivery_pace_tod_dow |> 
  ggplot(aes(x = dow, y = avg_daily_pace_tod_dow, fill = tod)) +
  geom_col(position = "dodge") +
  labs(
    title = "Freetown 2024: Delivery Pace by Day of Week & Time of Day",
    x = "Day of the Week", 
    y = "Average Deliveries per Enumerator", 
    fill = "Time of Day (3-hour brackets)"
  ) 
ggsave("freetown_2024_pace_dow_tod.png")

writeLines(strwrap(" 
Summary Characteristics: Filtered vs. Unfiltered Data

The Freetown 2024 data analysis showcases impressively consistent performance when comparing both filtered and unfiltered datasets. The mean daily deliveries stand at 66.48 for unfiltered data compared to 64.35 for filtered data, representing only a minimal 3.20% difference. The median performance figures of 66.37 deliveries for unfiltered data versus 65.02 for filtered data further reinforce this consistency, while the standard deviation of approximately 8.59 to 9.74 deliveries indicates moderate but acceptable variability in daily performance. When examining high-performance thresholds, the data reveals that the 80th percentile reaches 72 to 76 deliveries per day, while the 90th percentile extends to 74 or more deliveries, establishing clear benchmarks for exceptional operational days.

Temporal Performance Patterns

Day of Week Analysis

Peak Performance Days represent the operational sweet spot where multiple factors align to create optimal delivery conditions. Monday emerges as the standout performer with 73.73 average daily deliveries, establishing itself as the highest-performing day of the week. Tuesday follows closely with 71.77 average daily deliveries, followed by Wednesday which has a strong performance of 70.60 average daily deliveries. These three days collectively demonstrate the operation's capacity to handle increased volume while maintaining service quality.

Moderate Performance Days include Thursday with an average daily delivery pace of 67.61, followed closely by Saturday with 65.76 average daily deliveries, and finally Friday with 65.48 deliveries. These days likely represent periods where demand patterns, staffing levels, or logistical factors create more standard operational conditions.

Sunday is the lowest performance day with an average of 23.25 daily deliveries. This dramatic reduction is to be expected as enumerators should not be working on Sundays.

Time of Day Performance

The 12-14 hour period stands out as the peak performance window, achieving 29.13 deliveries and representing the most productive time slot of the operational day. The morning period from 09-11 hours follows with a strong performance of 27.29 average deliveries. The afternoon window from 15-17 hours follows with an average of 23.65 deliveries.

Lower Activity Periods include the night and early morning periods outside standard hours with an average of 10.02 deliveries followed by early morning periods from 06-08 hours with an average of 6.65 deliveries all of which is to be expected as enumerators should not be working these hours.

Visualizations

Distribution Analysis

This distribution pattern displays multiple performance clusters with clear peaks occurring around 60, 65, and 75 deliveries per day, indicating that average delivery pace evolved in distinct performance modes rather than following a simple linear progression as the distribution spans from approximately 40 deliveries up to 80 or more deliveries per day.

Temporal Trend Analysis

The time series analysis reveals significant day-to-day variability that tells a story of dynamic average delivery pace. The performance range extends dramatically from a low of 40 deliveries to a peak of 80 deliveries per day, showcasing the full spectrum of operational capability. Notable performance peaks occur throughout the operational period, where the operation achieves over 75 deliveries per day. The data also shows significant performance drops during various periods. However, there are consistent recovery patterns that follow these drops.", width = 80))
```

```{r}
#| label: Freetown 2023 Analysis
#| include: false

# Loading and cleaning the data: merging the 2 datasets for FCC 2023
fcc_2023 <- bind_rows(
  read_csv("fcc_2023_team_1-3.csv"),
  read_csv("fcc_2023_team_4-6.csv")
) |> 
  clean_names()

# Check for date parsing issues
bad_dates <- fcc_2023 |>  
  mutate(datetime = dmy_hms(time_delivery_date)) |> 
  filter(is.na(datetime) & !is.na(time_delivery_date)) |> 
  distinct(time_delivery_date)

# Clean up the time_delivery_date column to ensure formatting consistency
fcc_2023 <- fcc_2023 |> 
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    )
  )

# Split property_type column values into multiple rows (each counts as a delivery)
fcc_2023 <- fcc_2023 |> 
  separate_rows(property_type, sep = ",\\s*") 

# Transforming: unfiltered data (all deliveries)
fcc_2023_unfiltered <- fcc_2023 |> 
  filter(!is.na(time_delivery_date)) |> 
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    ),
    date = as_date(datetime)
  )|> 
  count(date, enteredby, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate(
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |> 
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Transforming: filtered data (accurate deliveries ≤ 0.08 km)
fcc_2023_filtered <- fcc_2023 |> 
  filter(time_gps_distance_original <= 0.08) |> 
  filter(!is.na(time_delivery_date)) |>
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    ),
    date = as_date(datetime)
  ) |> 
  count(date, enteredby, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate(
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |> 
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Summary statistics
freetown_stats_unfiltered <- fcc_2023_unfiltered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),
    median = median(daily_avg, na.rm = TRUE),
    sd     = sd(daily_avg, na.rm = TRUE),
    iqr    = IQR(daily_avg, na.rm = TRUE),
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)
  )
table_text <- capture.output(
  kable(freetown_stats_unfiltered, format = "simple", caption = "Freetown 2023 Unfiltered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2023_delivery_pace_unfiltered.txt")
View(freetown_stats_unfiltered)

freetown_stats_filtered <- fcc_2023_filtered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),
    median = median(daily_avg, na.rm = TRUE),
    sd     = sd(daily_avg, na.rm = TRUE),
    iqr    = IQR(daily_avg, na.rm = TRUE),
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)
  )
table_text <- capture.output(
  kable(freetown_stats_filtered, format = "simple", caption = "Freetown 2023 Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2023_delivery_pace_filtered.txt")
View(freetown_stats_filtered)

# Comparison
freetown_comparison_2023 <- tibble(
  unfiltered_mean = mean(fcc_2023_unfiltered$daily_avg, na.rm = TRUE),
  filtered_mean   = mean(fcc_2023_filtered$daily_avg, na.rm = TRUE)
) |> 
  mutate(
    diff_percent = (unfiltered_mean - filtered_mean) / unfiltered_mean * 100
  )
table_text <- capture.output(
  kable(freetown_comparison_2023, format = "simple", caption = "Freetown 2023 Unfiltered vs. Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2023_delivery_pace_comparison.txt")
View(freetown_comparison_2023)

# Time of day & day of week
fcc_2023_tod_dow <- fcc_2023 |> 
  filter(!is.na(time_delivery_date)) |>
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    ),
    date = as_date(datetime),
    dow = wday(date, label = TRUE, abbr = TRUE),
    hour = hour(datetime),
    tod = case_when(
      hour >= 6  & hour < 9  ~ "06–08",
      hour >= 9  & hour < 12 ~ "09–11",
      hour >= 12 & hour < 15 ~ "12–14",
      hour >= 15 & hour < 18 ~ "15–17",
      TRUE ~ "Other (Night/Early Morning)"
    )
  )

# Delivery pace by time of day, day of week
fcc_2023_delivery_pace_tod <- fcc_2023_tod_dow |> 
  count(date, tod, enteredby, name = "enum_deliv") |> 
  group_by(tod) |> 
  summarise(avg_daily_pace_tod = mean(enum_deliv), .groups = "drop")
table_text <- capture.output(
  kable(fcc_2023_delivery_pace_tod, format = "simple", caption = "Freetown 2023 Time of Day: Average Delivery Pace")
)
writeLines(table_text, "fcc_2023_delivery_pace_tod.txt")
View(fcc_2023_delivery_pace_tod)

fcc_2023_delivery_pace_dow <- fcc_2023_tod_dow |> 
  count(date, dow, enteredby, name = "enum_deliv") |> 
  group_by(dow) |> 
  summarise(avg_daily_pace_dow = mean(enum_deliv), .groups = "drop")
table_text <- capture.output(
  kable(fcc_2023_delivery_pace_dow, format = "simple", caption = "Freetown 2023 Day of Week: Average Delivery Pace")
)
writeLines(table_text, "fcc_2023_delivery_pace_dow.txt")
View(fcc_2023_delivery_pace_dow)

fcc_2023_delivery_pace_tod_dow <- fcc_2023_tod_dow |> 
  count(date, dow, tod, enteredby, name = "enum_deliv") |> 
  group_by(dow, tod) |> 
  summarise(avg_daily_pace_tod_dow = mean(enum_deliv), .groups = "drop")

# Visualizations
# (a) Time-series plot
fcc_2023_unfiltered |> 
  ggplot(aes(x = date, y = daily_avg)) +
  geom_line(color = "blue", linewidth = 1) +
  geom_point(size = 1.5, color = "black") +
  labs(
    title = "Freetown 2023: Average Accurate Daily Deliveries", 
    x = "Date", 
    y = "Average Accurate Daily Deliveries"
  ) 
ggsave("freetown_2023_timeseries.png")

# (b) Histogram + density curve
fcc_2023_unfiltered |> 
  ggplot(aes(x = daily_avg)) +
  geom_histogram(aes(y = after_stat(density)), 
                 bins = 15, fill = "lightblue", alpha = 0.6, color = "black") +
  geom_density(color = "darkblue", linewidth = 1) +
  labs(
    title = "Distribution of Daily Delivery Pace (Freetown 2023)", 
    x = "Deliveries per Day", 
    y = "Density"
  ) 
ggsave("freetown_2023_distribution.png")

# (c) Delivery pace by day of week & time of day
fcc_2023_delivery_pace_tod_dow |> 
  ggplot(aes(x = dow, y = avg_daily_pace_tod_dow, fill = tod)) +
  geom_col(position = "dodge") +
  labs(
    title = "Freetown 2023: Delivery Pace by Day of Week & Time of Day",
    x = "Day of the Week", 
    y = "Average Deliveries per Enumerator", 
    fill = "Time of Day (3-hour brackets)"
  ) 
ggsave("freetown_2023_pace_dow_tod.png")

writeLines(strwrap
("Summary Characteristics: Filtered vs. Unfiltered Data

The Freetown 2023 data analysis showcases good performance consistency when comparing both filtered and unfiltered datasets. The mean daily deliveries stand at 62.07 for unfiltered data compared to 59.08 for filtered data, representing a 4.82% difference. The median performance figures of 65.29 deliveries for unfiltered data versus 61.31 for filtered data further reinforce this pattern, while the standard deviation of approximately 12.46 to 12.78 deliveries indicates moderate variability in daily performance. When examining high-performance thresholds, the data reveals that the 80th percentile reaches 71 to 74 deliveries per day, while the 90th percentile extends to 73 or more deliveries, establishing clear benchmarks for exceptional operational days.

Temporal Performance Patterns

Day of Week Analysis

Peak Performance Days represent the operational sweet spot where multiple factors align to create optimal delivery conditions. Monday emerges as the standout performer with 68.97 average daily deliveries, establishing itself as the highest-performing day of the week. Wednesday follows closely with 68.21 average daily deliveries, followed by Saturday which has a strong performance of 67.54 average daily deliveries. These three days collectively demonstrate the operation's capacity to handle increased volume while maintaining service quality.
Moderate Performance Days include Tuesday with an average daily delivery pace of 67.37, followed closely by Friday with 65.80 average daily deliveries, and finally Thursday with 63.59 deliveries. These days likely represent periods where demand patterns, staffing levels, or logistical factors create more standard operational conditions.

Sunday is the lowest performance day with an average of 41.60 daily deliveries. This reduction represents a significant decrease from working day levels while maintaining substantial operational activity.

Time of Day Performance

The 12-14 hour period stands out as the peak performance window, achieving 28.63 deliveries and representing the most productive time slot of the operational day. The morning period from 09-11 hours follows with a strong performance of 25.90 average deliveries. The afternoon window from 15-17 hours follows with an average of 23.77 deliveries.

Lower Activity Periods include the night and early morning periods outside standard hours with an average of 10.42 deliveries followed by early morning periods from 06-08 hours with an average of 6.52 deliveries all of which is to be expected as enumerators should not be working these hours.

Visualizations

Distribution Analysis

This distribution pattern displays multiple performance clusters with clear peaks occurring around 55, 65, and 70 deliveries per day, indicating that average delivery pace evolved in distinct performance modes rather than following a simple linear progression as the distribution spans from approximately 35 deliveries up to 80 or more deliveries per day.

Temporal Trend Analysis

The time series analysis reveals significant day-to-day variability that tells a story of dynamic average delivery pace. The performance range extends dramatically from a low of 35 deliveries to a peak of 80 deliveries per day, showcasing the full spectrum of operational capability. Notable performance peaks occur throughout the operational period, where the operation achieves over 70 deliveries per day. The data also shows significant performance drops during various periods. However, there are consistent recovery patterns that follow these drops.", width = 80))


```

```{r}
#| label: Freetown 2022 Analysis
#| include: false

# Loading and cleaning the data: merging the 2 datasets for FCC 2022
fcc_2022 <- bind_rows(
  read_csv("fcc_2022_team_1-3.csv"),
  read_csv("fcc_2022_team_4-6.csv")
) |> 
  clean_names()

# Check which dates failed to parse
bad_dates <- fcc_2022 |>  
  mutate(datetime = dmy_hms(time_delivery_date)) |> 
  filter(is.na(datetime) & !is.na(time_delivery_date)) |> 
  distinct(time_delivery_date)

# Clean up the time_delivery_date column for uniform format
fcc_2022 <- fcc_2022 |> 
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    )
  )

# Split property_type column values into multiple rows
fcc_2022 <- fcc_2022 |> 
  separate_rows(property_type, sep = ",\\s*") 

# Transforming: unfiltered data (all deliveries)
fcc_2022_unfiltered <- fcc_2022 |> 
  filter(!is.na(time_delivery_date)) |> 
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    ),
    date = as_date(datetime)
  )|> 
  count(date, enteredby, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate(
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |> 
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Transforming: filtered data (accurate deliveries ≤ 0.08 km)
fcc_2022_filtered <- fcc_2022 |> 
  filter(time_gps_distance_original <= 0.08) |> 
  filter(!is.na(time_delivery_date)) |>
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    ),
    date = as_date(datetime)
  ) |> 
  count(date, enteredby, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate(
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |> 
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Summary statistics
freetown_stats_unfiltered <- fcc_2022_unfiltered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),
    median = median(daily_avg, na.rm = TRUE),
    sd     = sd(daily_avg, na.rm = TRUE),
    iqr    = IQR(daily_avg, na.rm = TRUE),
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)
  )
table_text <- capture.output(
  kable(freetown_stats_unfiltered, format = "simple", caption = "Freetown 2022 Unfiltered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2022_delivery_pace_unfiltered.txt")
View(freetown_stats_unfiltered)

freetown_stats_filtered <- fcc_2022_filtered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),
    median = median(daily_avg, na.rm = TRUE),
    sd     = sd(daily_avg, na.rm = TRUE),
    iqr    = IQR(daily_avg, na.rm = TRUE),
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)
  )
table_text <- capture.output(
  kable(freetown_stats_filtered, format = "simple", caption = "Freetown 2022 Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2022_delivery_pace_filtered.txt")
View(freetown_stats_filtered)

# Comparison
freetown_comparison_2022 <- tibble(
  unfiltered_mean = mean(fcc_2022_unfiltered$daily_avg, na.rm = TRUE),
  filtered_mean   = mean(fcc_2022_filtered$daily_avg, na.rm = TRUE)
) |> 
  mutate(
    diff_percent = (unfiltered_mean - filtered_mean) / unfiltered_mean * 100
  )
table_text <- capture.output(
  kable(freetown_comparison_2022, format = "simple", caption = "Freetown 2022 Unfiltered vs. Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2022_delivery_pace_comparison.txt")
View(freetown_comparison_2022)

# Time of day & day of week
fcc_2022_tod_dow <- fcc_2022 |> 
  filter(!is.na(time_delivery_date)) |>
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    ),
    date = as_date(datetime),
    dow = wday(date, label = TRUE, abbr = TRUE),
    hour = hour(datetime),
    tod = case_when(
      hour >= 6  & hour < 9  ~ "06–08",
      hour >= 9  & hour < 12 ~ "09–11",
      hour >= 12 & hour < 15 ~ "12–14",
      hour >= 15 & hour < 18 ~ "15–17",
      TRUE ~ "Other (Night/Early Morning)"
    )
  )

# Delivery pace by time of day, day of week
fcc_2022_delivery_pace_tod <- fcc_2022_tod_dow |> 
  count(date, tod, enteredby, name = "enum_deliv") |> 
  group_by(tod) |> 
  summarise(avg_daily_pace_tod = mean(enum_deliv), .groups = "drop")
table_text <- capture.output(
  kable(fcc_2022_delivery_pace_tod, format = "simple", caption = "Freetown 2022 Time of Day: Average Delivery Pace")
)
writeLines(table_text, "fcc_2022_delivery_pace_tod.txt")
View(fcc_2022_delivery_pace_tod)

fcc_2022_delivery_pace_dow <- fcc_2022_tod_dow |> 
  count(date, dow, enteredby, name = "enum_deliv") |> 
  group_by(dow) |> 
  summarise(avg_daily_pace_dow = mean(enum_deliv), .groups = "drop")
table_text <- capture.output(
  kable(fcc_2022_delivery_pace_dow, format = "simple", caption = "Freetown 2022 Day of Week: Average Delivery Pace")
)
writeLines(table_text, "fcc_2022_delivery_pace_dow.txt")
View(fcc_2022_delivery_pace_dow)

fcc_2022_delivery_pace_tod_dow <- fcc_2022_tod_dow |> 
  count(date, dow, tod, enteredby, name = "enum_deliv") |> 
  group_by(dow, tod) |> 
  summarise(avg_daily_pace_tod_dow = mean(enum_deliv), .groups = "drop")

# Visualizations
# (a) Time-series plot
fcc_2022_unfiltered |> 
  ggplot(aes(x = date, y = daily_avg)) +
  geom_line(color = "blue", linewidth = 1) +
  geom_point(size = 1.5, color = "black") +
  labs(
    title = "Freetown 2022: Average Accurate Daily Deliveries", 
    x = "Date", 
    y = "Average Accurate Daily Deliveries"
  ) 
ggsave("freetown_2022_timeseries.png")

# (b) Histogram + density curve
fcc_2022_unfiltered |> 
  ggplot(aes(x = daily_avg)) +
  geom_histogram(aes(y = after_stat(density)), 
                 bins = 15, fill = "lightblue", alpha = 0.6, color = "black") +
  geom_density(color = "darkblue", linewidth = 1) +
  labs(
    title = "Distribution of Daily Delivery Pace (Freetown 2022)", 
    x = "Deliveries per Day", 
    y = "Density"
  ) 
ggsave("freetown_2022_distribution.png")

# (c) Delivery pace by day of week & time of day
fcc_2022_delivery_pace_tod_dow |> 
  ggplot(aes(x = dow, y = avg_daily_pace_tod_dow, fill = tod)) +
  geom_col(position = "dodge") +
  labs(
    title = "Freetown 2022: Delivery Pace by Day of Week & Time of Day",
    x = "Day of the Week", 
    y = "Average Deliveries per Enumerator", 
    fill = "Time of Day (3-hour brackets)"
  ) 
ggsave("freetown_2022_pace_dow_tod.png")

writeLines(strwrap
("Summary Characteristics: Filtered vs. Unfiltered Data

The Freetown 2022 data analysis showcases moderate performance consistency when comparing both filtered and unfiltered datasets. The mean daily deliveries stand at 55.04 for unfiltered data compared to 51.04 for filtered data, representing a 7.27% difference. The median performance figures of 59.43 deliveries for unfiltered data versus 55.42 for filtered data further reinforce this pattern, while the standard deviation of approximately 10.29 to 10.65 deliveries indicates moderate variability in daily performance. When examining high-performance thresholds, the data reveals that the 80th percentile reaches 59 to 63 deliveries per day, while the 90th percentile extends to 61 or more deliveries, establishing clear benchmarks for exceptional operational days.

Temporal Performance Patterns

Day of Week Analysis

Peak Performance Days represent the operational sweet spot where multiple factors align to create optimal delivery conditions. Monday emerges as the standout performer with 59.86 average daily deliveries, establishing itself as the highest-performing day of the week. Tuesday follows closely with 59.48 average daily deliveries, followed by Thursday which has a strong performance of 59.08 average daily deliveries. These three days collectively demonstrate the operation's capacity to handle increased volume while maintaining service quality.

Moderate Performance Days include Wednesday with an average daily delivery pace of 59.05, followed closely by Friday with 57.88 average daily deliveries, and finally Saturday with 57.50 deliveries. These days likely represent periods where demand patterns, staffing levels, or logistical factors create more standard operational conditions.

Sunday is the lowest performance day with an average of 44.35 daily deliveries. This reduction represents a significant decrease from working day levels while maintaining substantial operational activity.

Time of Day Performance

The 12-14 hour period stands out as the peak performance window, achieving 24.93 deliveries and representing the most productive time slot of the operational day. The morning period from 09-11 hours follows with a strong performance of 21.81 average deliveries. The afternoon window from 15-17 hours follows with an average of 19.08 deliveries.

Lower Activity Periods include the night and early morning periods outside standard hours with an average of 7.91 deliveries followed by early morning periods from 06-08 hours with an average of 7.45 deliveries all of which is to be expected as enumerators should not be working these hours.

Visualizations

Distribution Analysis

This distribution pattern displays multiple performance clusters with clear peaks occurring around 45, 55, and 60 deliveries per day, indicating that average delivery pace evolved in distinct performance modes rather than following a simple linear progression as the distribution spans from approximately 30 deliveries up to 70 or more deliveries per day.

Temporal Trend Analysis

The time series analysis reveals significant day-to-day variability that tells a story of dynamic average delivery pace. The performance range extends dramatically from a low of 30 deliveries to a peak of 70 deliveries per day, showcasing the full spectrum of operational capability. Notable performance peaks occur throughout the operational period, where the operation achieves over 60 deliveries per day. The data also shows significant performance drops during various periods. However, there are consistent recovery patterns that follow these drops.", width = 80))


```

```{r}
#| label: Freetown 2021 Analysis
#| include: false

# Loading and cleaning the data: merging the 2 datasets for FCC 2021
fcc_2021 <- bind_rows(
  read_csv("fcc_2021_team_1-3.csv"),
  read_csv("fcc_2021_team_4-6.csv")
) |> 
  clean_names()

# Check which dates failed to parse
bad_dates <- fcc_2021 |>  
  mutate(datetime = dmy_hms(time_delivery_date)) |> 
  filter(is.na(datetime) & !is.na(time_delivery_date)) |> 
  distinct(time_delivery_date)

# Clean up the time_delivery_date column for uniform format
fcc_2021 <- fcc_2021 |> 
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    )
  )

# Split property_type column values into multiple rows
fcc_2021 <- fcc_2021 |> 
  separate_rows(property_type, sep = ",\\s*") 

# Transforming: unfiltered data (all deliveries)
fcc_2021_unfiltered <- fcc_2021 |> 
  filter(!is.na(time_delivery_date)) |> 
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    ),
    date = as_date(datetime)
  )|> 
  count(date, enteredby, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate(
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |> 
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Transforming: filtered data (accurate deliveries ≤ 0.08 km)
fcc_2021_filtered <- fcc_2021 |> 
  filter(time_gps_distance_original <= 0.08) |> 
  filter(!is.na(time_delivery_date)) |>
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    ),
    date = as_date(datetime)
  ) |> 
  count(date, enteredby, name = "enum_daily_deliv") |> 
  group_by(date) |> 
  summarise(daily_avg = mean(enum_daily_deliv), .groups = "drop") |> 
  mutate(
    Q1 = quantile(daily_avg, 0.25),
    Q3 = quantile(daily_avg, 0.75),
    IQR = Q3 - Q1
  ) |> 
  filter(
    daily_avg >= Q1 - 1.5 * IQR,
    daily_avg <= Q3 + 1.5 * IQR
  )

# Summary statistics
freetown_stats_unfiltered <- fcc_2021_unfiltered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),
    median = median(daily_avg, na.rm = TRUE),
    sd     = sd(daily_avg, na.rm = TRUE),
    iqr    = IQR(daily_avg, na.rm = TRUE),
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)
  )
table_text <- capture.output(
  kable(freetown_stats_unfiltered, format = "simple", caption = "Freetown 2021 Unfiltered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2021_delivery_pace_unfiltered.txt")
View(freetown_stats_unfiltered)

freetown_stats_filtered <- fcc_2021_filtered |> 
  summarise(
    mean   = mean(daily_avg, na.rm = TRUE),
    median = median(daily_avg, na.rm = TRUE),
    sd     = sd(daily_avg, na.rm = TRUE),
    iqr    = IQR(daily_avg, na.rm = TRUE),
    p80    = quantile(daily_avg, 0.80, na.rm = TRUE),
    p90    = quantile(daily_avg, 0.90, na.rm = TRUE)
  )
table_text <- capture.output(
  kable(freetown_stats_filtered, format = "simple", caption = "Freetown 2021 Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2021_delivery_pace_filtered.txt")
View(freetown_stats_filtered)

# Comparison
freetown_comparison_2021 <- tibble(
  unfiltered_mean = mean(fcc_2021_unfiltered$daily_avg, na.rm = TRUE),
  filtered_mean   = mean(fcc_2021_filtered$daily_avg, na.rm = TRUE)
) |> 
  mutate(
    diff_percent = (unfiltered_mean - filtered_mean) / unfiltered_mean * 100
  )
table_text <- capture.output(
  kable(freetown_comparison_2021, format = "simple", caption = "Freetown 2021 Unfiltered vs. Filtered Data: Average Delivery Pace")
)
writeLines(table_text, "fcc_2021_delivery_pace_comparison.txt")
View(freetown_comparison_2021)

# Time of day & day of week
fcc_2021_tod_dow <- fcc_2021 |> 
  filter(!is.na(time_delivery_date)) |>
  mutate(
    datetime = parse_date_time(
      time_delivery_date,
      orders = c("dmy HMS", "mdy HMS", "dmy HM", "mdy HM")
    ),
    date = as_date(datetime),
    dow = wday(date, label = TRUE, abbr = TRUE),
    hour = hour(datetime),
    tod = case_when(
      hour >= 6  & hour < 9  ~ "06–08",
      hour >= 9  & hour < 12 ~ "09–11",
      hour >= 12 & hour < 15 ~ "12–14",
      hour >= 15 & hour < 18 ~ "15–17",
      TRUE ~ "Other (Night/Early Morning)"
    )
  )

# Delivery pace by time of day, day of week
fcc_2021_delivery_pace_tod <- fcc_2021_tod_dow |> 
  count(date, tod, enteredby, name = "enum_deliv") |> 
  group_by(tod) |> 
  summarise(avg_daily_pace_tod = mean(enum_deliv), .groups = "drop")
table_text <- capture.output(
  kable(fcc_2021_delivery_pace_tod, format = "simple", caption = "Freetown 2021 Time of Day: Average Delivery Pace")
)
writeLines(table_text, "fcc_2021_delivery_pace_tod.txt")
View(fcc_2021_delivery_pace_tod)

fcc_2021_delivery_pace_dow <- fcc_2021_tod_dow |> 
  count(date, dow, enteredby, name = "enum_deliv") |> 
  group_by(dow) |> 
  summarise(avg_daily_pace_dow = mean(enum_deliv), .groups = "drop")
table_text <- capture.output(
  kable(fcc_2021_delivery_pace_dow, format = "simple", caption = "Freetown 2021 Day of Week: Average Delivery Pace")
)
writeLines(table_text, "fcc_2021_delivery_pace_dow.txt")
View(fcc_2021_delivery_pace_dow)

fcc_2021_delivery_pace_tod_dow <- fcc_2021_tod_dow |> 
  count(date, dow, tod, enteredby, name = "enum_deliv") |> 
  group_by(dow, tod) |> 
  summarise(avg_daily_pace_tod_dow = mean(enum_deliv), .groups = "drop")

# Visualizations
# (a) Time-series plot
fcc_2021_unfiltered |> 
  ggplot(aes(x = date, y = daily_avg)) +
  geom_line(color = "blue", linewidth = 1) +
  geom_point(size = 1.5, color = "black") +
  labs(
    title = "Freetown 2021: Average Accurate Daily Deliveries", 
    x = "Date", 
    y = "Average Accurate Daily Deliveries"
  ) 
ggsave("freetown_2021_timeseries.png")

# (b) Histogram + density curve
fcc_2021_unfiltered |> 
  ggplot(aes(x = daily_avg)) +
  geom_histogram(aes(y = after_stat(density)), 
                 bins = 15, fill = "lightblue", alpha = 0.6, color = "black") +
  geom_density(color = "darkblue", linewidth = 1) +
  labs(
    title = "Distribution of Daily Delivery Pace (Freetown 2021)", 
    x = "Deliveries per Day", 
    y = "Density"
  ) 
ggsave("freetown_2021_distribution.png")

# (c) Delivery pace by day of week & time of day
fcc_2021_delivery_pace_tod_dow |> 
  ggplot(aes(x = dow, y = avg_daily_pace_tod_dow, fill = tod)) +
  geom_col(position = "dodge") +
  labs(
    title = "Freetown 2021: Delivery Pace by Day of Week & Time of Day",
    x = "Day of the Week", 
    y = "Average Deliveries per Enumerator", 
    fill = "Time of Day (3-hour brackets)"
  ) 
ggsave("freetown_2021_pace_dow_tod.png")

writeLines(strwrap
("Summary Characteristics: Filtered vs. Unfiltered Data

The Freetown 2021 data analysis showcases good performance consistency when comparing both filtered and unfiltered datasets. The mean daily deliveries stand at 59.67 for unfiltered data compared to 56.41 for filtered data, representing a 5.47% difference. The median performance figures of 62.56 deliveries for unfiltered data versus 59.13 for filtered data further reinforce this pattern, while the standard deviation of approximately 12.28 to 13.00 deliveries indicates moderate variability in daily performance. When examining high-performance thresholds, the data reveals that the 80th percentile reaches 64 to 68 deliveries per day, while the 90th percentile extends to 70 or more deliveries, establishing clear benchmarks for exceptional operational days.

Temporal Performance Patterns

Day of Week Analysis

Peak Performance Days represent the operational sweet spot where multiple factors align to create optimal delivery conditions. Friday emerges as the standout performer with 64.71 average daily deliveries, establishing itself as the highest-performing day of the week. Wednesday follows closely with 62.98 average daily deliveries, followed by Thursday which has a strong performance of 61.24 average daily deliveries. These three days collectively demonstrate the operation's capacity to handle increased volume while maintaining service quality.

Moderate Performance Days include Saturday with an average daily delivery pace of 60.06, followed closely by Tuesday with 59.47 average daily deliveries, and finally Monday with 58.42 deliveries. These days likely represent periods where demand patterns, staffing levels, or logistical factors create more standard operational conditions.

Sunday is the lowest performance day with an average of 53.36 daily deliveries. This reduction represents a moderate decrease from working day levels while maintaining substantial operational activity.

Time of Day Performance

The 12-14 hour period stands out as the peak performance window, achieving 24.10 deliveries and representing the most productive time slot of the operational day. The morning period from 09-11 hours follows with a strong performance of 20.29 average deliveries. The afternoon window from 15-17 hours follows with an average of 19.90 deliveries.

Lower Activity Periods include the night and early morning periods outside standard hours with an average of 8.23 deliveries followed by early morning periods from 06-08 hours with an average of 6.49 deliveries all of which is to be expected as enumerators should not be working these hours.

Visualizations

Distribution Analysis

This distribution pattern displays multiple performance clusters with clear peaks occurring around 50, 60, and 65 deliveries per day, indicating that average delivery pace evolved in distinct performance modes rather than following a simple linear progression as the distribution spans from approximately 20 deliveries up to 80 or more deliveries per day.

Temporal Trend Analysis

The time series analysis reveals significant day-to-day variability that tells a story of dynamic average delivery pace. The performance range extends dramatically from a low of 20 deliveries to a peak of 80 deliveries per day, showcasing the full spectrum of operational capability. Notable performance peaks occur throughout the operational period, where the operation achieves over 75 deliveries per day. The data also shows significant performance drops during various periods. However, there are consistent recovery patterns that follow these drops.", width = 80))

```

# **Results & Discussion**

## **Filtered vs. Unfiltered Data Comparison**

This section evaluates how filtering for accurate deliveries affects the average delivery daily pace and highlights trends over multiple years for Makeni, Kenema, and Freetown.

```{r, fig.width=20, fig.height=12, dpi=300}
#| label: Combined Analysis - Filtered vs Unfiltered
#| echo: false

# Adding city & year to unfiltered datasets
makeni_2025_unfiltered   <- mcc_2025_unfiltered   |> mutate(city = "Makeni", year = 2025, dataset = "Unfiltered")
kenema_2025_unfiltered   <- kcc_2025_unfiltered   |> mutate(city = "Kenema", year = 2025, dataset = "Unfiltered")
kenema_2024_unfiltered   <- kcc_2024_unfiltered   |> mutate(city = "Kenema", year = 2024, dataset = "Unfiltered")
freetown_2025_unfiltered <- fcc_2025_unfiltered   |> mutate(city = "Freetown", year = 2025, dataset = "Unfiltered")
freetown_2024_unfiltered <- fcc_2024_unfiltered   |> mutate(city = "Freetown", year = 2024, dataset = "Unfiltered")
freetown_2023_unfiltered <- fcc_2023_unfiltered   |> mutate(city = "Freetown", year = 2023, dataset = "Unfiltered")
freetown_2022_unfiltered <- fcc_2022_unfiltered   |> mutate(city = "Freetown", year = 2022, dataset = "Unfiltered")
freetown_2021_unfiltered <- fcc_2021_unfiltered   |> mutate(city = "Freetown", year = 2021, dataset = "Unfiltered")

# Adding city & year to filtered datasets
makeni_2025_filtered   <- mcc_2025_filtered   |> mutate(city = "Makeni", year = 2025, dataset = "Filtered")
kenema_2025_filtered   <- kcc_2025_filtered   |> mutate(city = "Kenema", year = 2025, dataset = "Filtered")
kenema_2024_filtered   <- kcc_2024_filtered   |> mutate(city = "Kenema", year = 2024, dataset = "Filtered")
freetown_2025_filtered <- fcc_2025_filtered   |> mutate(city = "Freetown", year = 2025, dataset = "Filtered")
freetown_2024_filtered <- fcc_2024_filtered   |> mutate(city = "Freetown", year = 2024, dataset = "Filtered")
freetown_2023_filtered <- fcc_2023_filtered   |> mutate(city = "Freetown", year = 2023, dataset = "Filtered")
freetown_2022_filtered <- fcc_2022_filtered   |> mutate(city = "Freetown", year = 2022, dataset = "Filtered")
freetown_2021_filtered <- fcc_2021_filtered   |> mutate(city = "Freetown", year = 2021, dataset = "Filtered")

# Combining all datasets
all_deliveries <- bind_rows(
  makeni_2025_filtered, kenema_2025_filtered, kenema_2024_filtered,
  freetown_2025_filtered, freetown_2024_filtered, freetown_2023_filtered, 
  freetown_2022_filtered, freetown_2021_filtered, makeni_2025_unfiltered, 
  kenema_2025_unfiltered, kenema_2024_unfiltered, freetown_2025_unfiltered, 
  freetown_2024_unfiltered, freetown_2023_unfiltered, freetown_2022_unfiltered, 
  freetown_2021_unfiltered
)

# Summary characteristics by city, year, dataset
summary_all <- all_deliveries |>
  group_by(city, year, dataset) |>
  summarise(
    mean_daily = mean(daily_avg, na.rm = TRUE),
    median_daily = median(daily_avg, na.rm = TRUE),
    p80 = quantile(daily_avg, 0.8, na.rm = TRUE),
    p90 = quantile(daily_avg, 0.9, na.rm = TRUE),
    .groups = "drop"
  )

summary_change <- summary_all |> # adding year-over-year % changes in average daily deliveries (ignores NAs for first year)
  arrange(city, dataset, year) |>  
  group_by(city, dataset) |>
  mutate(
    pct_change = (mean_daily - lag(mean_daily)) / lag(mean_daily) * 100
  ) |>
  ungroup()
kable( summary_change, caption = "Year-over-Year Percentage Change in Average Daily Deliveries" )


avg_change <- summary_change |>  # adding total average % change per city overtime
  group_by(city, dataset) |> 
  summarise( avg_pct_change = mean(pct_change, na.rm = TRUE), 
             .groups = "drop" 
  )
kable( avg_change, caption = "Average Percentage Change in Delivery Pace Over Time by City" )


# Visualizations
# (a) Summary bar chart
summary_all |>
  ggplot(aes(x = factor(year), y = mean_daily, fill = dataset)) +
  geom_col(position = "dodge") +
  facet_wrap(~city) +
  labs(
    title = "Average Daily Deliveries: Filtered vs Unfiltered",
    x = "Year",
    y = "Mean Deliveries per Day",
    fill = "Dataset Type"
  ) +
  theme_minimal(base_size = 20) +  # <-- sets global text size
  theme(
    plot.title = element_text(size = 30, face = "bold"),
    axis.title = element_text(size = 26),
    axis.text  = element_text(size = 22),
    legend.title = element_text(size = 26),
    legend.text  = element_text(size = 22),
    strip.text = element_text(size = 20, face = "bold")
  )
ggsave("filtered_vs_unfiltered_all_cities_bar.png", width = 16, height = 10, dpi = 300)

# (b) Distribution: filtered vs unfiltered
all_deliveries |>
  ggplot(aes(x = daily_avg, fill = dataset)) +
  geom_histogram(aes(y = after_stat(density)), bins = 20, alpha = 0.5, position = "identity") +
  geom_density(aes(color = dataset), linewidth = 1) +
  facet_grid(year ~ city) +
  labs(
    title = "Distribution of Daily Delivery Pace: Filtered vs Unfiltered",
    x = "Deliveries per Day",
    y = "Density",
    fill = "Dataset Type",
    color = "Dataset Type"
  )+
  theme_minimal(base_size = 20) +  # <-- sets global text size
  theme(
    plot.title = element_text(size = 30, face = "bold"),
    axis.title = element_text(size = 26),
    axis.text  = element_text(size = 22),
    legend.title = element_text(size = 26),
    legend.text  = element_text(size = 22),
    strip.text = element_text(size = 20, face = "bold")
  )
ggsave("filtered_vs_unfiltered_all_cities_distribution.png", width = 30, height = 24, dpi = 300)



```

#### Makeni (2025)

In Makeni, the analysis shows that the mean daily deliveries were 31.68 in the unfiltered dataset compared to 31.28 in the filtered dataset, representing a minimal 1.24% difference. The median values were 33.09 and 32.60 deliveries, respectively, while the standard deviation ranged from 14.6 to 14.8 deliveries, indicating moderate but acceptable variability. High-performance benchmarks, as reflected in the 80th and 90th percentiles, reached 45–46 and 48+ deliveries per day, respectively. These findings indicate that filtering has minimal impact on overall performance, suggesting that errors or anomalies in the dataset do not meaningfully distort operational trends.

The distribution analysis confirms a normal distribution with a peak around 30–35 deliveries, with filtered and unfiltered datasets closely overlapping. The narrower delivery range (approximately 15–50 deliveries) with concentration around 25–40 deliveries supports stable but moderate-capacity operations. Overall, Makeni demonstrates stable average daily deliveries, with consistent average output and predictable peaks.

#### Kenema (2024-2025)

In Kenema, the 2024–2025 comparison shows that the mean daily deliveries increased slightly from 38.83 in the filtered dataset in 2024 to 40.15 in 2025, representing a 3.40% increase, while the unfiltered dataset shows a comparable 4.49% increase over the same period. The median daily deliveries remained similar, though the standard deviation decreased sharply from approximately 21.5 in 2024 to 9.0 in 2025, reflecting a marked improvement in consistency and reduced day-to-day variability. The 80th and 90th percentiles indicate that high-output days became more predictable, even as extreme peaks were slightly lower. These trends suggest that Kenema’s operations in 2025 were smoother and more reliable compared to 2024. Overall, filtering has a minimal effect on the assessment of operational performance but ensures that the analysis reflects accurate, error-free data.

The distribution analysis shows that Kenema exhibits a substantial transformation between its 2024 and 2025 distributions. In 2024, the distribution was bimodal, with peaks around 20 and 45 deliveries and a broad performance range spanning 0–80 deliveries. This pattern reflected inconsistency, with significant low-performance clustering around 0–20 deliveries. By 2025, operations stabilized significantly. The distribution transformed into a concentrated single-peaked normal pattern centered around 40 deliveries, with a narrower range of 15–60 deliveries. Filtering impact remained low, with filtered and unfiltered distributions closely aligned. The transformation indicates that Kenema’s operations became more consistent, eliminating extreme low-performance days and supporting predictable scheduling overtime.

#### Freetown (2021-2025)

In Freetown, analysis across multiple years demonstrates both growth and fluctuations in operational performance. The mean daily deliveries in filtered dataset were 56.41 in 2021, 51.04 in 2022, 59.08 in 2023, 64.35 in 2024, and 59.74 in 2025. In unfiltered dataset, the mean deliveries were slightly higher each year, ranging from 59.67 in 2021 to 64.30 in 2025. The median and high-performance thresholds followed similar trends, reflecting growing peak capacity over time.

The year-over-year performance changes provide additional insight. In the filtered dataset, Freetown experienced a decrease of 9.52% from 2021 to 2022, followed by strong growth of 15.75% in 2022–2023 and 8.93% in 2023–2024, before declining 7.17% in 2024–2025. Unfiltered data shows slightly smaller fluctuations, with increases of 12.77% from 2022–2023 and 7.11% from 2023–2024, followed by a decrease of 3.28% in 2024–2025. The average annual change rates indicate gradual long-term growth, with 1.997% for filtered data and 2.212% for unfiltered data.

These results indicate that, while daily averages fluctuate, Freetown has achieved a general upward trend in operational capacity over the years. Filtering reduces the impact of anomalies on the data but does not materially change the observed trends, suggesting that the improvements and declines reflect true operational performance rather than delivery errors.

Freetown demonstrates consistently high delivery volumes and generally normal distribution shapes across all years. The distributions were predominantly single-peaked and normal, with minor right skew in 2021. Filtered and unfiltered dataset closely overlapped in all years, though small differences in tail regions. Year-over-year, Freetown shows clear distribution evolution. In 2021–2022, the distribution shifted slightly left, reflecting a modest decrease in performance. From 2022–2025, distributions shifted gradually rightward, demonstrating improvements in daily delivery performance, with peak activity consistently concentrated in 45–75 deliveries. The mode consistently aligns with the mean, confirming symmetric, normal distributions, and standard deviations remained moderate, supporting predictable delivery patterns.

#### Cross-City Insights

Across cities, several patterns emerge. Freetown demonstrates gradual long-term growth, with increasing mean deliveries and higher high-performance thresholds over time, though fluctuations in certain years indicate periods of operational adjustment or demand variation. Kenema shows measurable improvements in 2025 compared to 2024, with higher average deliveries and significantly reduced variability, indicating more consistent and reliable daily operations. Makeni, while only available for 2025, shows stable performance with predictable peaks and minimal difference between filtered and unfiltered data.

Overall, filtered data confirms that inaccurate deliveries do not significantly distort average daily delivery pace, while unfiltered data reflects the full operational reality. Across all cities, the differences between filtered and unfiltered datasets are small, reinforcing that the analysis of mean, median, and peak performance provides a reliable picture of delivery operations. Year-over-year and average annual change rates highlight the evolution of capacity, with Kenema and Freetown showing positive growth trends and improved operational stability.

## **Time of Day & Day of Week Analysis**

This section analyzes average delivery pace by **day of week** and **time of day**, highlighting cross-year and cross-city trends.

```{r, fig.width=50, fig.height=40, dpi=300}
#| label: Combined Analysis - Time and Day Patterns
#| echo: false

# Visualization: Delivery pace by day of week and time of day (all cities, all years)

all_delivery_pace_tod_dow <- bind_rows(
  fcc_2021_delivery_pace_tod_dow|> mutate(city = "Freetown", year = 2021),
  fcc_2022_delivery_pace_tod_dow |> mutate(city = "Freetown", year = 2022),
  fcc_2023_delivery_pace_tod_dow |> mutate(city = "Freetown", year = 2023),
  fcc_2024_delivery_pace_tod_dow |> mutate(city = "Freetown", year = 2024),
  fcc_2025_delivery_pace_tod_dow |> mutate(city = "Freetown", year = 2025),
  kcc_2024_delivery_pace_tod_dow |> mutate(city = "Kenema", year = 2024),
  kcc_2025_delivery_pace_tod_dow |> mutate(city = "Kenema", year = 2025),
  mcc_2025_delivery_pace_tod_dow |> mutate(city = "Makeni", year = 2025)
)

all_delivery_pace_tod_dow |> 
  ggplot(aes(x = dow, y = avg_daily_pace_tod_dow, fill = tod)) +
  geom_col(position = "dodge") +
  facet_grid(city ~ year) +
  scale_x_discrete(labels = c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")) +
  labs(
    title = "Delivery Pace by Day of Week & Time of Day (All Cities, All Years)",
    x = "Day of the Week",
    y = "Average Deliveries per Enumerator",
    fill = "Time of Day (3-hour brackets)"
  ) +
  theme_minimal(base_size = 30) +  # <-- lighter theme, bigger fonts
  theme(
    plot.title = element_text(size = 60, face = "bold"),
    axis.title = element_text(size = 40),
    axis.text.x = element_text(size = 32, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 32),
    legend.title = element_text(size = 50),
    legend.text  = element_text(size = 40),
    strip.text   = element_text(size = 30, face = "bold")
  )
  
ggsave("tod_dow_all_cities_bar.png", width = 40, height = 30, dpi = 300)


# Time of Day Analysis
all_delivery_pace_tod <- bind_rows(
  fcc_2021_delivery_pace_tod|> mutate(city = "Freetown", year = 2021),
  fcc_2022_delivery_pace_tod |> mutate(city = "Freetown", year = 2022),
  fcc_2023_delivery_pace_tod |> mutate(city = "Freetown", year = 2023),
  fcc_2024_delivery_pace_tod |> mutate(city = "Freetown", year = 2024),
  fcc_2025_delivery_pace_tod |> mutate(city = "Freetown", year = 2025),
  kcc_2024_delivery_pace_tod |> mutate(city = "Kenema", year = 2024),
  kcc_2025_delivery_pace_tod |> mutate(city = "Kenema", year = 2025),
  mcc_2025_delivery_pace_tod |> mutate(city = "Makeni", year = 2025)
)

detailed_pace_tod <- all_delivery_pace_tod |> # detailed averages of time of day by city and year
  filter(!is.na(avg_daily_pace_tod)) |>
  select(city, year, tod, avg_daily_pace_tod) |>
  arrange(city, year, tod)

city_pace_tod <- all_delivery_pace_tod |> # city-level averages by time of day (across years for each city)
  filter(!is.na(avg_daily_pace_tod)) |>
  group_by(city, tod) |>
  summarise(
    city_avg_pace_tod = mean(avg_daily_pace_tod, na.rm = TRUE),
    years_count = n_distinct(year),
    min_pace = min(avg_daily_pace_tod, na.rm = TRUE),
    max_pace = max(avg_daily_pace_tod, na.rm = TRUE),
    .groups = "drop"
  )
kable(city_pace_tod, format = "simple", caption = "Average Delivery Pace by Time of Day Across All Cities and Years")
View(city_pace_tod)

# Day of Week analysis

all_delivery_pace_dow <- bind_rows(
  fcc_2021_delivery_pace_dow|> mutate(city = "Freetown", year = 2021),
  fcc_2022_delivery_pace_dow |> mutate(city = "Freetown", year = 2022),
  fcc_2023_delivery_pace_dow |> mutate(city = "Freetown", year = 2023),
  fcc_2024_delivery_pace_dow |> mutate(city = "Freetown", year = 2024),
  fcc_2025_delivery_pace_dow |> mutate(city = "Freetown", year = 2025),
  kcc_2024_delivery_pace_dow |> mutate(city = "Kenema", year = 2024),
  kcc_2025_delivery_pace_dow |> mutate(city = "Kenema", year = 2025),
  mcc_2025_delivery_pace_dow |> mutate(city = "Makeni", year = 2025)
)


detailed_pace_dow <- all_delivery_pace_dow |> # detailed averages of time of day by city and year
  filter(!is.na(avg_daily_pace_dow)) |>
  select(city, year, dow, avg_daily_pace_dow) |>
  arrange(city, year, dow)

city_pace_dow <- all_delivery_pace_dow |> # city-level averages by time of day (across years for each city)
  filter(!is.na(avg_daily_pace_dow)) |>
  group_by(city, dow) |>
  summarise(
    city_avg_pace_dow = mean(avg_daily_pace_dow, na.rm = TRUE),
    years_count = n_distinct(year),
    min_pace = min(avg_daily_pace_dow, na.rm = TRUE),
    max_pace = max(avg_daily_pace_dow, na.rm = TRUE),
    .groups = "drop"
  )
kable(city_pace_dow, format = "simple", caption = "Average Delivery Pace by Time of Day Across All Cities and Years")
View(city_pace_dow)




```

### Makeni 

#### 2025 Performance

In Makeni's single year of observation (2025), the highest-performing day is Tuesday, averaging 48.66 deliveries per enumerator, followed by Monday at 43.57 and Friday at 36.30. Wednesday and Thursday are moderate-performance days with averages of 29.22 and 27.44 deliveries respectively, while Saturday delivers 23.41 on average. Sunday represents the lowest activity at 13.80 deliveries, likely reflecting reduced operational hours or staffing on this traditional rest day.

Examining the time-of-day performance in 2025, the 12:00–14:00 period emerges as the most productive window at 18.39 deliveries per enumerator, followed by 09:00–11:00 with 14.94 deliveries, and 15:00–17:00 with 12.89 deliveries. Early morning (06:00–08:00) and night/early morning periods remain predictably low at 5.91 and 4.97 deliveries respectively.

#### Average Performance

Since only one year of data is available, the 2025 figures represent Makeni's averages. The combination of day-of-week and time-of-day analysis indicates that Makeni's peak operational capacity is concentrated in early-to-mid-week and midday periods, particularly Tuesday midday hours. The significant drop-off from Tuesday's peak to other weekdays (up to 73% lower on Sunday) suggests potential opportunities for more consistent resource allocation and operational standardization. Without multi-year data, it remains unclear whether these patterns represent stable operational norms or a single-year snapshot that may not reflect longer-term trends.

## Kenema 

#### 2024 Performance

In 2024, Kenema's operations showed a distinctive pattern with Thursday and Monday as the highest-performing days. Thursday averaged 61.79 deliveries per enumerator, and Monday averaged 61.43 deliveries. Tuesday performed moderately at 43.01 deliveries, while Friday averaged 43.66 deliveries. Wednesday and Saturday were notably weak, averaging just 21.03 and 21.74 deliveries respectively—representing a 65-66% decline from the peak days. There are no recorded deliveries on Sundays for Kenema's 2024 data.

Time-of-day patterns in 2024 showed the 12:00–14:00 period leading at 21.24 deliveries per enumerator, followed by 09:00–11:00 at 20.87 deliveries and 15:00–17:00 at 15.02 deliveries. Early morning (06:00–08:00) averaged 5.97 deliveries, while night/early morning periods averaged 6.72 deliveries.

#### 2025 Performance

In 2025, Kenema's operational patterns shifted significantly. Sunday emerged as an important day with 49.50 deliveries per enumerator. Wednesday rebounded dramatically to 46.36 deliveries—more than doubling from its 2024 low. Thursday remained relatively strong at 45.70 deliveries, though down from 2024's peak. Monday declined to 38.80 deliveries (a 37% drop from 2024), Tuesday fell to 37.73 deliveries, and Friday decreased to 36.34 deliveries. Saturday remained the lowest weekday at 33.52 deliveries but improved 54% from 2024.

Time-of-day performance in 2025 showed some compression. The 12:00–14:00 period averaged 20.55 deliveries per enumerator (slightly lower than 2024), 09:00–11:00 dropped to 18.00 deliveries (14% decline), and 15:00–17:00 remained stable at 14.71 deliveries. Early morning (06:00–08:00) decreased to 4.87 deliveries, while night/early morning periods fell to 4.73 deliveries.

#### Average Performance (2024-2025)

Across both years, Kenema averages reveal moderate overall performance with high variability. Monday averages 50.11 deliveries per enumerator across the two years (though this masks the 37% year-over-year decline). Thursday averages 53.74 deliveries—the highest across both years. Tuesday averages 40.37 deliveries, Wednesday averages 33.69 deliveries (despite the 121% improvement from 2024 to 2025), Friday averages 40.00 deliveries, and Saturday averages 27.63 deliveries. Sunday's single-year figure of 49.50 cannot be averaged meaningfully.

For time-of-day averages, the 12:00–14:00 window averages 20.90 deliveries per enumerator across both years, 09:00–11:00 averages 19.44 deliveries, 15:00–17:00 averages 14.87 deliveries, 06:00–08:00 averages 5.42 deliveries, and night/early morning periods average 5.72 deliveries.

The dramatic year-over-year variations in day-of-week performance (up to 121% increases or 37% decreases) illlustrates a significant operational restructuring from 2024-2025 delivery in Kenema. The relative stability in time-of-day patterns, by contrast, indicates more consistent intra-day scheduling practices even as weekly allocation changes.

## Freetown 

#### 2021 Performance

In 2021, Freetown's weekday performance showed Friday leading at 64.71 deliveries per enumerator, followed by Wednesday at 62.98, Thursday at 61.24, Monday at 61.62, Tuesday at 59.47, and Saturday at 59.98. Sunday recorded 37.50 deliveries.

Time-of-day performance in 2021 showed the 12:00–14:00 period at 24.10 deliveries per enumerator, 09:00–11:00 at 20.29 deliveries, 15:00–17:00 at 19.08 deliveries, 06:00–08:00 at 6.49 deliveries, and night/early morning at 7.91 deliveries.

#### 2022 Performance

In 2022, weekday patterns shifted slightly with Monday rising to 63.72 deliveries per enumerator, Tuesday at 64.77, Wednesday at 64.04, Thursday at 59.08, Friday at 57.88, and Saturday at 57.50. Sunday improved notably to 44.35 deliveries.

Time-of-day performance in 2022 showed the 12:00–14:00 period at 26.26 deliveries, 09:00–11:00 at 23.43 deliveries, 15:00–17:00 at 20.16 deliveries, 06:00–08:00 at 6.70 deliveries, and night/early morning at 8.62 deliveries.

#### 2023 Performance

In 2023, Monday reached 64.62 deliveries per enumerator, Tuesday peaked at 71.77, Wednesday at 70.60, Thursday at 67.73, Friday at 68.32, and Saturday at 69.69. Sunday increased to 53.36 deliveries—the highest Sunday performance across all years.

Time-of-day performance in 2023 showed the 12:00–14:00 period at 28.20 deliveries, 09:00–11:00 at 25.04 deliveries, 15:00–17:00 at 23.85 deliveries, 06:00–08:00 at 8.67 deliveries, and night/early morning at 13.94 deliveries.

#### 2024 Performance

In 2024, Monday delivered 58.42 deliveries per enumerator, Tuesday 60.35, Wednesday 59.05, Thursday 60.74, Friday 61.78, and Saturday 64.65. Sunday dropped sharply to 23.25 deliveries—the lowest Sunday performance across all years.

Time-of-day performance in 2024 showed the 12:00–14:00 period at 27.02 deliveries, 09:00–11:00 at 24.57 deliveries, 15:00–17:00 at 21.04 deliveries, 06:00–08:00 at 6.95 deliveries, and night/early morning at 10.11 deliveries.

#### 2025 Performance

In 2025, Monday surged to 73.73 deliveries per enumerator—the highest single day-of-week performance across all years and cities. Tuesday reached 66.42, Wednesday 70.65, Thursday 70.49, Friday 68.49, and Saturday 69.08. Sunday data is not provided for 2025.

Time-of-day performance in 2025 showed the 12:00–14:00 period at 30.88 deliveries—the highest midday performance across all years. The 09:00–11:00 period reached 27.29 deliveries, 15:00–17:00 at 21.09 deliveries, 06:00–08:00 at 7.12 deliveries, and night/early morning at 10.02 deliveries.

#### Average Performance (2021-2025)

Across five years, Freetown's averages demonstrate consistently high productivity with moderate variation. Monday averages 65.23 deliveries per enumerator (ranging from 58.42 to 73.73—a 26% spread), Tuesday averages 64.55 deliveries (59.47 to 71.77—a 21% spread), Wednesday averages 65.26 deliveries (59.05 to 70.65—a 20% spread), Thursday averages 63.85 deliveries (59.08 to 67.73—a 15% spread), Friday averages 64.44 deliveries (57.88 to 68.32—an 18% spread), and Saturday averages 64.11 deliveries (57.50 to 69.69—a 21% spread). Sunday averages 39.78 deliveries across the four years with data (23.25 to 53.36—a dramatic 129% spread indicating evolving weekend operation policies).

Time-of-day averages across the five years show the 12:00–14:00 period at 27.53 deliveries per enumerator (24.10 to 30.88—a 28% increase trajectory), 09:00–11:00 at 24.45 deliveries (20.29 to 27.29—a 35% increase trajectory), 15:00–17:00 at 22.05 deliveries (19.08 to 23.85—a 25% increase trajectory), 06:00–08:00 at 7.16 deliveries (6.49 to 8.67—a 34% spread), and night/early morning at 10.10 deliveries (7.91 to 13.94—a 76% spread).

The year-over-year trajectory shows Freetown's operations strengthening over time, with 2025 representing peak performance in most metrics. The relative stability in weekday averages (Monday through Saturday showing less than 3% variation from each other) demonstrates exceptional operational consistency. The gradual improvements in midday productivity (28% increase from 2021 to 2025) suggest successful operational optimization and capacity building over the five-year period.

#### Cross-City Insights

When comparing cities, several consistent patterns emerge. Firstly, mid-day hours (12:00–14:00) consistently represent the most productive window across all cities and years. Morning (09:00–11:00) and afternoon (15:00–17:00) periods complement these peaks, enabling near-continuous operational productivity. Whiile early morning (06:00–08:00) and night shifts consistently show low activity, which is to be expected given operational policies restricting work during these hours.

Mid-week or early-week days generally deliver the highest volume, while Sundays consistently record the lowest deliveries.

# **Conclusions & Policy Recommendations**

The analysis reveals distinct delivery pace profiles across the three cities, with each demonstrating unique strengths and improvement opportunities. Freetown demonstrates the highest average daily deliveries with an average of 59.74 daily deliveries, though it experiences some performance fluctuations that require careful management. Kenema shows the most encouraging improvement trajectory, increasing from 38.83 deliveries in 2024 to 40.15 deliveries in 2025 while significantly reducing day-to-day variability. Makeni maintains stable and predictable performance with 31.28 daily deliveries, providing a solid foundation for sustainable growth. These findings inform targeted recommendations designed to set sustainable daily delivery paces across all three cities in the 2025-2026 fiscal year and beyond.

## Proposed City-Specific Delivery Pace Targets for the 2025–2026 Fiscal Year

Targets are established as ranges, providing both lower and upper benchmarks to guide enumerator performance. The minimum threshold functions as a safeguard to ensure consistent productivity across the team, discouraging underperformance and reinforcing accountability. It also serves as a critical reference point for the design of delivery zones, ensuring workloads are distributed fairly and remain manageable for all enumerators. Conversely, the maximum threshold is intended to recognize and reward the contributions of higher-achieving enumerators, while at the same time preventing risks associated with excessive workloads, such as fatigue, mistakes, or declining data quality. This approach strikes a balance between motivating steady improvement, accommodating varying capacities, and maintaining high performance standards without compromising quality.

#### Makeni

The 2025 unfiltered data for Makeni indicates an average daily delivery pace of 31.68 deliveries per day, establishing the foundation for 2026 target calculations. Given the absence of multi-year historical data for Makeni, Kenema's average annual delivery pace growth rate of 4.49% was used as a proxy metric. This approach is statistically justified given that both cities operate at similar scales and demonstrate comparable performance characteristics, making Kenema's growth pattern a reasonable predictive indicator for Makeni's potential development tra

Applying the 4.49% growth adjustment to Makeni's baseline performance yields a projected 2026 baseline target of 33.10 deliveries per day. The high-performance threshold analysis utilizes Makeni's 90th percentile performance level of 48 deliveries per day from 2025 operations. When adjusted for the same growth trajectory, this high-performance threshold projects to 50.16 deliveries per day for 2026 operations.

**Recommended 2026 target delivery range: 33-50 deliveries per day.**

#### Kenema

Kenema's 2025 unfiltered data has a mean of 40.96 deliveries per day. Projecting forward with the established growth trajectory of 4.49%, Kenema's 2026 baseline target reaches 42.80 deliveries per day. The high-performance threshold analysis begins with Kenema's demonstrated capacity of 51.55 deliveries per day during peak operational periods. Applying the growth adjustment yields a high-performance target of 53.86 deliveries per day.

**Recommended 2026 target delivery range: 43-54 deliveries per day.**

#### Freetown

Freetown's delivery operations represent the most established and highest-capacity center among the three locations, with 2024 baseline performance averaging 66.48 deliveries per day. The city's multi-year performance analysis yields a conservative annual growth rate of 2.21%, reflecting the mature nature of operations and the challenges associated with scaling already high-volume activities. Applying Freetown's growth trajectory projects a 2026 baseline target of 69.41 deliveries per day. High-performance threshold analysis utilizes Freetown's demonstrated peak capacity of 78.38 deliveries per day in 2024, representing the city's ability to handle exceptional demand periods. With growth adjustments applied, the high-performance target for 2026 reaches 81.90 deliveries per day, establishing an ambitious but achievable ceiling for exceptional performance.

\*Note: 2024 is used as a baseline for Freetown given the data anomalies of the 2025 Freetown data.

**Recommended 2026 target delivery range: 69-82 deliveries per day.**

## Policy Considerations For Future Delivery Pace Target Setting

Establishing delivery pace targets is not an exact science. While data provides essential benchmarks, effective target-setting must incorporate broader practical considerations, including operational realities and staff input.

-   **Field consultation:** Feedback from field staff is critical to ensure that targets remain realistic, achievable, and reflective of actual working conditions.
-   **Ongoing monitoring:** Targets should be informed by continuous tracking of annual changes in average daily delivery pace, with attention to percentage growth patterns. This ensures projections are aligned with both potential and capacity.

Importantly, there will come a point where adopting a **fixed delivery pace target**—rather than pursuing continual increases—becomes the most sustainable and effective approach. This stage has not yet been reached in Kenema and Makeni, where post–property tax reform implementation still presents opportunities for growth. However, in more mature systems such as Freetown’s, conditions for fixed targets may emerge sooner. This is important because while incremental increases can encourage progress, targets that exceed realistic capacity risk undermining morale if perceived as unattainable. Higher unrealistic delivery pace targets and compounded growth rates—even as modest as 4–5% annually—can surpass actual infrastructure or staffing improvements prompting errors, rework, and inefficiencies, and negating intended productivity gains.

#### Indicators for Maintaining a Fixed Rate

-   If year-to-year growth levels off and significant gains are no longer observed, maintaining a fixed baseline may be more meaningful than ongoing adjustments.
-   If filtered and unfiltered data show minimal divergence, indicating fewer outliers and a stable system, a fixed target can help consolidate consistent performance.
-   If year-over-year increases in average daily delivery pace fall below 2%.
-   If delivery pace distributions stabilize, with little upward movement in median or peak values.
-   If signs of operational strain emerge (e.g., rising error rates, absenteeism, or delays).
-   If staff consistently meet but do not exceed targets despite additional training or support.
